### 2015-06-26 Rmd/magisterial-district-ABM-select.Rmd

---
title: "Districts for ABM"
author: "Lyndon Estes"
date: "June 26, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

## Overview

Selection of magisterial districts and associated landcover sets for use in ABM example.

## Data
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# landcover data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*07sum_mask*")))
gti <- raster(paths)  # gti 2007 rather than 2011, to match census data

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2007))
lc_list <- gti - dlist_1km

# magisterial district data cropland statistics (2011, but only to look at for 
# selection 
load("external/ext_data/gti_md.rda")
mu <- data.frame(t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(mu) <- c("N", "na", "f")

# district shapes, filtered by N missing data, minimum fraction, and area
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
md@data[mdsel, 7:8]
mdsel <- as.numeric(row.names(mdfilt))

# Examine, including in QGIS
plot(md)
plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
         layer = "md_abms", driver = "SQLite")

```

### Selection

Looking at the data in QGIS, including some rainfall maps from earlier papers and the inter-tubes, it looks like the following districts are a good match:

   + Clocolan (269)
   + Ficksburg (274)
   + Fouriesburg (275)
   + Marquard (288)

```{r, eval=FALSE}
abm_md <- md[mdsel[mdsel %in% c(269, 274:275, 288)], ]
writeOGR(abm_md, dsn = "external/ext_data/md4.sqlite", layer = "md4", 
         driver = "SQLite")

# Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
# interest's sake
gti_abm <- extract(gti, abm_md)
abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(abm_mu) <- c("N", "na", "f")
abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
 # -1.0  0.0 -0.4 -0.2  # tinf differences!
```

### Extraction
```{r, eval=FALSE}
gti4 <- crop(gti, extent(abm_md))
lc4 <- crop(lc_list, extent(abm_md))
cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif")
```

### Visuals
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(raster)
library(rgdal)
library(lmisc)

setwd(full_path(proj_root("SAcropland"), "SAcropland"))
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
abm_md <- readOGR("external/ext_data/md4.sqlite", layer = "md4")
cb <- brick("external/ext_data/cropland-md4.tif")
```

```{r}
par(mar = c(0, 0, 0, 0))
plot(md, col = "grey", border = "transparent")
plot(abm_md, col = "red", border = "grey30", add = TRUE)
```

The four selected districts.

```{r}
par(mar = c(0, 0, 0, 0))
plot(cb[[1]], axes = FALSE, box = FALSE)
plot(abm_md, add = TRUE)
```

2007 cropland fraction in the four districts.

```{r}
par(mar = c(0, 0, 0, 0), mfrow = c(2, 2))
for(i in 2:5) {
  plot(cb[[i]], axes = FALSE, box = FALSE, zlim = c(0, 101))
  plot(abm_md, add = TRUE)
}
```

Cropland fractions estimated by South Africa's landcover product, GLobCover, MODIS, and Geo-Wiki for the same 4 districts. Note the substantial variability.