### 2015-07-06 Rmd/carbon-bias.Rmd
Initial versions of carbon bias estimation complete. Added utility functions used in various analyses to library. Minor edit to magisterial-district-bias.Rmd

---
title: "Carbon bias"
author: "Lyndon Estes"
date: "June 24, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

## Bias in carbon stock estimates
Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.

## Data

### Paths
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)
library(data.table)
library(SAcropland)
library(RColorBrewer)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")
p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
```

### Carbon values

Starting with the carbon look-up tables used for Africa by Ruesch & Gibbs (2008), inputs downloaded from [here](http://cdiac.ornl.gov/epubs/ndp/global_carbon/carbon_documentation.html#methods).
```{r, eval=FALSE}
cfiles <- dir(p_carb, pattern = "m6", full.names = TRUE)
nms <- gsub("*.*m6|\\.txt", "", cfiles)
cval_list <- lapply(1:length(cfiles), function(x) {
  tab <- read.table(cfiles[x])
  tab <- tab[ c(1, 3)]
  colnames(tab) <- c("CL", nms[x])
  tab
})

# merge into single carbon table
mergefun <- function(x, y) merge(x, y, by = "CL", all.x = TRUE, all.y = TRUE)
ctab <- Reduce(mergefun, cval_list)

# Read in lookup table key
key <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[25:44]
key2 <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[47:57]
mtch <- sapply(gregexpr("[0-9]", key2), max)
lcs <- sapply(1:length(key2), function(x) substr(key2[x], 1, mtch[x]))
lcs <- gsub(" & ", ",", gsub("-", ":", lcs))

# Reshape and reduce according to carbon classes
ctab2 <- t(data.frame(sapply(lcs, function(x) {
  ind <- eval(parse(text = paste0("c(", x, ")")))
  as.numeric(as.vector(
    round(colMeans(ctab[ctab$CL %in% ind, -1], na.rm = TRUE))))
})))
rownames(ctab2) <- 1:nrow(ctab2)
colnames(ctab2) <- nms
ctab2 <- ctab2 * 0.01  # convert to tons/ha from 1000 kg/ha
ctab2 <- cbind(transform(lcs), ctab2)
colnames(ctab2)[1] <- "class"
```
<a href="#top">Back to top</a>

### Further compress classes that have the same carbon value.

  + 1:3, 6:8 => 1 (broadleaf and mixed forests)
  + 4:5 => Drop
  + 9, 10; 17 => 2 (Secondary forests, forest/cropland mosaic)
  + 11, 12, 15 => 3 (shrublands)
  + 20:23 => drop (water, snow, artificial surfaces)
  + 19 => drop (bare areas)

```{r, eval=FALSE}
ctabf <- round(rbind(colMeans(ctab2[c(1, 3), 2:ncol(ctab2)], na.rm = TRUE), 
                     colMeans(ctab2[10:11, 2:ncol(ctab2)], na.rm = TRUE), 
                     ctab2[4:6, 2:ncol(ctab2)]))

# Read in ecofloristic regions and calculate their areas for Africa
ecoflora <- readOGR("external/ext_data/africa_ecofloristic_zones.sqlite", 
                    layer = "africa_ecofloristic_zones")
ecoflora@proj4string <- CRS("+proj=longlat +datum=WGS84 +no_defs")
afalb <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 ", 
                "+y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
ecoflora_alb <- spTransform(ecoflora, CRS = CRS(afalb))
ecoflora_num <- cbind.data.frame(unique(ecoflora_alb@data),  
                                 "ID" = c("06", "08", "09", "15", "NA", "18",
                                          "20", "16", "17", "WA", "19", "07"),
                                 stringsAsFactors = FALSE)
ecoflora_alb$ID <- rep("NA", nrow(ecoflora_alb))
head(ecoflora_alb)
ecoflora_alb$ID <- ecoflora_num[match(ecoflora_alb$gez_term, 
                                      ecoflora_num$gez_term), "ID"]
# ecoflora_alb@data <- sp::merge(ecoflora_alb@data, ecoflora_num, 
#                                by = "gez_term", sort = FALSE, keep = TRUE)
ecoflora_alb@data[sample(1:nrow(ecoflora_alb@data), 2), ]
ecoflora_alb$area <- round(rgeos::gArea(ecoflora_alb, byid = TRUE) / 1000000, 1)
ecoareas <- sapply(ecoflora_num$ID, function(x) {
  sum(ecoflora_alb@data[ecoflora_alb$ID == x, "area"])
})
ecoareas <- ecoareas[!names(ecoareas) %in% c("WA", "NA")]
ecowgts <- ecoareas / sum(ecoareas)
ecowgts <- ecowgts[sort(names(ecowgts))]

# par(mar = rep(0, 4))
# plot(ecoflora_alb)
# plot(ecoflora_alb[ecoflora_alb$ID == "17", ], col = "red", add = TRUE)

i <- 1:ncol(ctabf)
ctabf$mu <- round(sapply(1:nrow(ctabf), function(x) {
  sum(ctabf[x, ] * ecowgts, na.rm = TRUE)
}), 1)
ctabf <- cbind(ctabf, t(apply(ctabf[, i], 1, function(x) range(x, na.rm=TRUE))))
colnames(ctabf)[(ncol(ctabf) - 1):ncol(ctabf)] <- c("min", "max")
rownames(ctabf) <- 1:nrow(ctabf)
LC <- c("forest", "second", "shrub", "grass", "sparse")
ctabo <- data.frame(t(ctabf[, c("mu", "min", "max")]))  # reclass table
colnames(ctabo) <- LC

```
<a href="#top">Back to top</a>

## Analysis
### Prepare raster data
```{r, eval=FALSE}
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
plot(lc_list[[1]])
plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- c(5, 10, 25, 50, 100)
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# need NA mask for weighting of aggregated value
namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
namask2 <- !is.na(namask)  # set NAs to zero
area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
gti_bins <- lapply(gti_agg, function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})

```
<a href="#top">Back to top</a>

### Create cropland carbon estimates
```{r, eval=FALSE}
# Apply carbon estimates
cc <- ctab2[ctab2$class == 16, 2]
carbon <- function(fcrop, cropC, noncropC) {
  carb <- fcrop * cropC + noncropC * (1 - fcrop)
  return(round(carb, 2))
}

cc <- ctab2[ctab2$class == 16, 2]
gti_c <- lapply(gti_agg, function(x) {
  r <- x$gti
  s <- stack(lapply(1:ncol(ctabo), function(y) {
    carbon(r, cc, ctabo[1, y])
  }))
  names(s) <- colnames(ctabo)
  s
})

lc_c <- lapply(lc_agg, function(x) {
  lc <- lapply(x, function(j) {
    s <- stack(lapply(1:ncol(ctabo), function(y) {
      carbon(j, cc, ctabo[1, y])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(lc) <- names(x)
  lc
})

# checks--right rasters being referenced? 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  cellStats(lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]] - 
              carbon(lc_agg[[tst[i, 1]]][[tst[i, 2]]], cc, ctabo[1, tst[i, 3]]),
              sum)
}) # all zeroes

```
<a href="#top">Back to top</a>

### Difference the carbon datasets
```{r, eval=FALSE}
# percent difference
pct_diff <- function(x, y) (x - y) / x * 100
c_pct_diff <- lapply(1:length(gti_c), function(x) {
  dif <- lapply(1:length(lc_c[[x]]), function(y) {
    s <- stack(lapply(1:nlayers(lc_c[[x]][[y]]), function(z) {
      p <- pct_diff(gti_c[[x]][[z]], lc_c[[x]][[y]][[z]])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(dif) <- names(lc_c[[x]])
  dif
})
names(c_pct_diff) <- names(lc_c)

# checks - right rasters being referenced 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  x <- gti_c[[tst[i, 1]]][[tst[i, 3]]]
  y <- lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  z <- c_pct_diff[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  cellStats(z - ((x - y) / x * 100), sum)
})  # zeroes

# Then for a map plot figure, calculate the mean pixel-wise percent difference
c_pct_diff_mu <- lapply(c_pct_diff, function(x) {
  lapply(x, function(y) calc(y, mean))
})

# # disggregate selected rasters at selected levels for plotting
namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
lev <- names(c_pct_diff_mu) 
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- c_pct_diff_mu[[y]][[x]]
    } else {
      r <- raster::disaggregate(c_pct_diff_mu[[y]][[x]], 
                                fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms

stats <- lapply(disagg, function(x) {
  sapply(x, function(y) {
    c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
  })
})
```
<a href="#top">Back to top</a>

## Outputs
### Carbon bias maps
```{r, eval=FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape

# Plotting colors
lims <- c(ceiling(min(sapply(stats, function(x) x[3, ]))), 
          floor(max(sapply(stats, function(x) x[21, ]))))
rng <- range(sapply(stats, range))
brks <- c(rng[1], lims[1], -45, -20, -10, -5, -1, 1, 5, 10, lims[2], rng[2])
#n_cols <- length(brks) - 1
colsall <- brewer.pal(n = 11, "Spectral")
cols <- c(colsall[1:6], "grey80", colsall[c(7, 9, 10, 11)])
#colorRampPalette(c("red", "tan", "grey80"))
#cols2 <- colorRampPalette(c("grey80", "green4", "blue"))
#cols <- c(cols1(8), c(cols2(5)))

# brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
# n_cols <- length(brks) - 1
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(disagg[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "carbon_bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```
<a href="#top">Back to top</a>

### Carbon bias statistics
#### Calculate weighted bias statistics
```{r, eval=FALSE}
# Helper functions to pass into data.table
# not used - switch on if quantiles needs
# bfn <- function(x, y) {
#   box_stats(x, weighted = TRUE, weight.opts = list("weights" = y))
# }
# bfna <- function(x, y) {
#   box_stats(abs(x), weighted = TRUE, weight.opts = list("weights" = y))
# }
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# check constancy of < 1/2% cropland being excluded
areas <- sapply(lc_agg, function(x) res(x[[1]])[1]^2 / 10000)
plot(areas, (areas * 0.005))  # scales
(areas * 0.005)  # 1/2 to 5000 ha

lev_vec <- names(c_pct_diff)
# created mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lc_union <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lc_union) <- lev_vec
# plot(lc_union$f50$globmu)
# plot(gti_agg$f50$gti)

# selection variables
nms1 <- c(colnames(ctabo), "wgt")
nms2 <- colnames(ctabo)
# bvals <- names(box_stats(sample(1:100, 200, replace = TRUE)))
bvals <- "mu"

dang <- Sys.time()
  #print(paste("..", x))
cb_stats <- lapply(lev_vec, function(x) {  # level
  #x <- lev_vec[4]
  print(paste(".", x))
  #levr <- names(c_pct_diff[[x]])
  l1 <- lapply(names(c_pct_diff[[x]]), function(y) {
    #y <- levr[1]
    print(paste("...", y))
    lc <- c_pct_diff[[x]][[y]]
    lcmask <- lc_union[[x]][[y]]

    # stack raster bins, cropland bins, and landcover set
    s <- stack(list("bin" = gti_bins[[x]], "wgt" = area_wgts[[x]][[1]], 
                    "mask" = lc_union[[x]][[y]], c_pct_diff[[x]][[y]]))
    DT <- na.omit(agroEcoTradeoff::as.data.table.raster(s))
    setkey(DT, "bin")
    # indexing columns
    #front1 <- data.table("bin" = rep("all", 6), "N" = rep(nrow(DT), 6))
    #front2 <- data.table("bin" = rep("abs", 6), "N" = rep(nrow(DT), 6))
    # stats, overall means
    #m <- cbind(bvals, DT[mask == 1, lapply(.SD, bfn, wgt), .SDcols = nms1])
    #ma <- cbind(bvals, DT[mask == 1, lapply(.SD, bfna, wgt), .SDcols = nms1])
    # stats per bin
    #b <- cbind(bvals, 
    #           DT[mask == 1, lapply(.SD, bfn, wgt), by = bin, .SDcols = nms1])
    #ba <- cbind(bvals, 
    #            DT[mask == 1, lapply(.SD, bfna, wgt), by = bin, .SDcols = nms1])    
    # output
    #odt <- rbind(cbind(front1, mu), cbind(front2, muabs), 
    #             binl[bin][, bin := paste0("a", bin)], 
    #             binl[binabs][, bin := paste0("b", bin)])[, wgt := NULL]

    fr1 <- data.table("bvals" = c("m", "m0", "ma", "ma0"))
    fr2 <- data.table("bin" = "all", "N" = nrow(DT))
    binl <- DT[, .N, by = bin]  # n obs per bin
    a <- round(rbind(DT[mask == 1, lapply(.SD, wm, wgt), .SDcols = nms1], 
                     DT[, lapply(.SD, wm, wgt), .SDcols = nms1], 
                     DT[mask == 1, lapply(.SD, wma, wgt), .SDcols = nms1],
                     DT[, lapply(.SD, wma, wgt), .SDcols = nms1]), 2)
    dtl <- list(DT[mask == 1, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1],
                DT[, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1], 
                DT[mask == 1, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1], 
                DT[, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1])
    b <- rbindlist(lapply(1:4, function(x) cbind(fr1[x], round(dtl[[x]], 2))))
    setkey(b, "bin")
    odt <- rbind(cbind(fr1, rbind(cbind(fr2, a))), binl[b])
  })
  named_out(l1, names(c_pct_diff[[x]]))
})      
dut <- Sys.time() - dang  # 3.65 minutes (vs 33 in earlier incarnation), 
# 24 sec for means only
names(cb_stats) <- lev_vec
#save(cb_stats, file = "external/ext_data/carbon_bias_tables.rda")
save(cb_stats, file = "external/ext_data/carbon_bias_tables_mus.rda")

extract_stat <- function(namevec, stats, i1, i2) {
  estats <- do.call(rbindlist(lapply(namevec, function(i) {
    data.table(sapply(stats, function(x) {
      unlist(sapply(x, function(y) {
        v <- y[bin == i1 & bvals == i2, i, with = FALSE]
        if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
        v
      }))
    }))
  }))
  out <- cbind(do.call(rbind, strsplit(rownames(estats), "\\.")), 
               data.table(round(estats, 2)))
  setnames(out, old = names(mu), new = )
  return(out)
}
N <- function(namevec, stats, i1, i2) {
  estats <- data.table(do.call(rbind, lapply(namevec, function(i) {
    o <- sapply(stats, function(x) {
      sapply(x, function(y) y[bin == i1 & bvals == i2, N][1])
    })
  })))
  return(estats)
}
```
<a href="#top">Back to top</a>

#### Plot means against scale, supplemental version 
```{r, eval=FALSE}
a <- extract_stat(LC, cb_stats, "all", "m")  # mean abs crop only
b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(a, c(2:1, 3:ncol(a)))
mu <- rbind(a, b)

a <- extract_stat(LC, cb_stats, "all", "ma")  # mean abs crop only
b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mua <- rbind(a, b)

a <- extract_stat(LC, cb_stats, "all", "m0")  # mean abs crop only
b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mu0 <- rbind(a, b)

a <- extract_stat(LC, cb_stats, "all", "ma0")  # mean abs crop only
b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mu0a <- rbind(a, b)

#yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
# yl <- range(all_mus[, 3:8, with = FALSE])

cols <- c("red", "orange3", "green4", "blue")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
lw <- 1.5

yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
yl <- round(yl / 10) * 10
yax <- seq(yl[1], yl[2], 10)
xax <- seq(1, 6, 5 / (length(yax) - 1))

# Version 1 of mean carbon bias against scale - probably for supplementals, if 
# retained
pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 5)
par(mfrow = c(3, 2), mar = c(1, 0, 1, 1.25), mgp = c(1, 0.5, 0), tcl = -0.3,  
    oma = c(6, 4, 1, 0))
for(f in 1:length(LC)) {
  plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
       ylab = "", xlab = "", 
       panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
  #grid(NA, NULL, lwd = 1, lty = 1)
  mtext(side = 3, text = pnms[f], line = 0)
  if(f == 5) axis(1, at = 1:6, labels = c(1, fact))
  if(f %in% c(1, 3, 5)) axis(2, at = yax, labels = yax, las = 2)
  if(f %in% c(2, 4)) axis(2, at = yax, labels = rep("", length(yax)), las = 2)
  for(i in 1:4) {
    lines(1:6, mu[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
          col = cols[i], lty = 2, lwd = lw)
  }
  for(i in 1:4) {
    lines(1:6, mua[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
          col = cols[i], lty = 1, lwd = lw)
  }
}
plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "", 
     panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
for(i in 1:4) {
 lines(1:6, all_mu[V1 == snms[i], lev_vec, with = FALSE], col = cols[i], 
       lty = 2, lwd = lw)
}
for(i in 1:4) {
  lines(1:6, all_mua[V1 == snms[i], lev_vec, with = FALSE], col = cols[i],
        lty = 1, lwd = lw)
}
mtext(side = 3, text = "Vegetation mean", line = 0)
axis(1, at = 1:6, labels = c(1, fact))
axis(2, at = yax, labels = rep("", length(yax)), las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 1.5)
mtext(side = 2, text = "% bias", outer = TRUE, line = 2)
par(xpd = NA)
x <- grconvertX(0.7, from = "ndc", to = "user")
y <- grconvertY(0.075, from = "ndc", to = "user")
legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = c(2, 1), 
       lwd = 2, bty = "n", cex = 1.2)
x <- grconvertX(0.1, from = "ndc", to = "user")
y <- grconvertY(0.11, from = "ndc", to = "user")
legend(x = x, y = y, legend = mcap, lty = 1, col = cols, lwd = 2, bty = "n",
       cex = 1.2)
dev.off()

# Statistics from non-croplands areas included in means, for supplementals
nmrepl <- cbind(snms, lcnms)
mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
mu0[, V1 := nmrepl[match(V1, nmrepl), 2]]
mu0a[, V1 := nmrepl[match(V1, nmrepl), 2]]
mu0s <- rbind(cbind(stat = "actual", mu0), cbind(stat = "absolute", mu0a))
mu0s_xtab <- xtable::xtable(mu0s)
print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex")

# Statistics from non-croplands areas included in means, for supplementals
nmrepl <- cbind(snms, lcnms)
mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
mu[, V1 := nmrepl[match(V1, nmrepl), 2]]
mua[, V1 := nmrepl[match(V1, nmrepl), 2]]
mus <- rbind(cbind(stat = "actual", mu), cbind(stat = "absolute", mua))
mus_xtab <- xtable::xtable(mus)
print(mu0s_xtab, type = "latex", 
      file = "paper/figures/C-bias-mean-cropland-SA.tex")

```
<a href="#top">Back to top</a>

#### Main text version of mean carbon bias against scale
```{r, eval=FALSE}
alph <- c(225, 75)
#Ccols <- brewer.pal(n = 7, name = "Greys")
LC2 <- c(LC[-grep("second|grass", LC)], "all")
# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
#x + xo[j] + o[k]
xa <- sapply(x, function(x) x + xo)
mutype <- c("mua", "mu")

o <- c(0, w)
pchs <- c("*", "+", "o", "-")
cx <- c(1, 0.7, 0.6, 1)
g1 <- "grey90"
# pchShow()
# ctabo
xl <- c(-0.5, 18)
yl <- c(-60, 70)
shd <- c(4.5, 10.5, 16.5)

pdf(full_path(p_fig, "carbon_veg_scale2.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
grid(NA, 13, lwd = 1, lty = 1)
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
for(i in 1:length(lev_vec)) {
  for(j in 1:length(snms)) {
    for(k in 1:length(mutype)) {
      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
      for(z in 1:length(LC2)) {
        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
               pch = pchs[z], type = "pt")
      }
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "% bias", outer = TRUE, line = 1)
legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex = 0.8)
text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
dev.off()

```
<a href="#top">Back to top</a>

```{r, eval=FALSE, echo=FALSE}
# Plot means against cropland density (to possibly be run - for now, 
# too much information)
mub <- do.call(rbind, lapply(1:20, function(x) {
  cbind("bin" = x, extract_stat(LC, cb_stats, x, "m"))
}))
muba <- do.call(rbind, lapply(1:20, function(x) {
  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma"))
}))
n <- do.call(rbind, lapply(1:20, function(x) {
  cbind("bin" = x, N(LC, cb_stats, x, "m"))
}))
#n[, mean(f50), by = bin]

pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
                       seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, 
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                         rep("grey80", 10)),
           "yl" = list(c(-100, 100), c(0, 100)), 
           "yl2" = c(-100, 100), "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
                            " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           "ycol" = c("grey70"))  # plot parameters

yax <- seq(-100, 100, 10)
xax <- seq(0, 22, 20 / (length(yax) - 2))

pdf(full_path(p_fig, "carbon_biases_1-100km_v1.pdf"), width = 7.5, height = 7)
par(mfrow = c(6, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,  
    oma = c(6, 4, 1, 0))
for(f in 1:length(LC)) {
  for(j in c(1, 4:6)) {#j <- 1; f <- 1
      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
       ylab = "", xlab = "", 
       panel.first = grid(NA, length(yax) - 1, lwd = 1, lty = 1))
      lines(pp$x, pp$y, lwd = 1)
    for(i in 1:4) {
      mub[V1 == snms[i] & V2 == LC[f], 
          lines(get(lev_vec[j]), col = cols[i], lty = 2, lwd = lw)]
      muba[V1 == snms[i] & V2 == LC[f], 
           lines(get(lev_vec[j]), col = cols[i], lty = 1, lwd = lw)]
    }
  }
}
dev.off()

ys <- c(-100, 100)
yax <- seq(ys[1], ys[2], 10)
xax <- seq(0, 22, 20 / (length(yax) - 2))

#yl <- list(c(-100, 100), c(0, 100))
yl <- seq(yax[1], yax[length(yax)], 10)
yll <- as.character(yl)
yll[seq(2, length(yll) - 1, 2)] <- ""
Ccols <- brewer.pal(n = 6, name = "Dark2") 
xaxl = ifelse(seq(5, 100, 5) %in% c(5, 25, 50, 75, 100), seq(5, 100, 5), " ")
x <- mub[V1 == snms[f] & V2 == LC[j], get(lev_vec[i])]
y <- 100
toohigh <- function(x, y, col = "black", yadj = 0, cex = 1.1) {
  h <- which(abs(x) > abs(y))
  if(length(h) > 0) {
    xi <- h[which.max(h)]
    text(xi, y = y + yadj, labels = round(x[xi]), cex = cex, col = col, 
         xpd = NA)
  }
}

pdf(full_path(p_fig, "carbon_biases_1-100km_v2.pdf"), width = 7.5, height = 7)
par(mfrow = c(4, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,  
    oma = c(6, 4, 1, 0))
for(f in 1:length(snms)) {
  for(i in c(1, 4:6)) {#i <- 1; j <- 1; f <- 2
      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
           ylab = "", xlab = "", 
           panel.first = grid(NA, (length(yax) - 1) / 2, lwd = 1, lty = 1))
      lines(pp$x, pp$y, lwd = 1)
    for(j in 1:length(LC)) {
      e <- parse(text = "V1 == snms[f] & V2 == LC[j]")
      mub[eval(e), 
          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]

      mub[V1 == snms[f] & V2 == LC[j], 
          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]
      mub[V1 == snms[f] & V2 == LC[j], toohigh(get(lev_vec[i]), ys[1])]
      muba[V1 == snms[f] & V2 == LC[j], 
           lines(get(lev_vec[i]), col = Ccols[j], lty = 1, lwd = lw)]
      axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
           tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
      axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
           tcl = -0.3, pos = pp$yl2[1])
      if(f == length(snms)) {
        axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
             tcl = -0.3, pos = pp$yl2[1])
      }
      if(i == 1) axis(2, at = yl, labels = yll, las = 2, mgp = c(1, 0.6, 0))
      if(i == 1 & j == 1) {
        mtext(lcnms[f], side = 2, line = 2.8, cex = 1, col = cols[f])
      }
    }
  }
}
dev.off()

i1 <- "sa30"; i2 <- "forest"; s1 <- "f1"
mub[V1 == i1 & V2 == i2, ]
mub[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
mubin[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
```

```{r, eval=FALSE, echo=FALSE}
# Checks on data values (embedded code)
# Carbon values and carbon biases - random selection of cells from c_tab_diff
# compared to recalculated values for those points
i <- c_pct_diff
for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
                c("sa30", "f50", "sparse"))) {
  x <- chk[1] #names(i[[1]])[3]
  y <- chk[2] #names(i)[4]
  tp <- chk[3]
  r <- i[[y]][[x]][[tp]]
  rid <- which(values(Which(r != 0)))
  sid <- sample(rid, 20)
  v1 <- r[sid]  # value to check
  sav <- gti_agg[[y]]$gti[sid]
  lcv <- lc_agg[[y]][[x]][sid]
  c1 <- carbon(sav, cropC = 5, noncropC = ctabo[1, tp])
  c2 <- carbon(lcv, cropC = 5, noncropC = ctabo[1, tp])
  #round((sav * 5) + (ctabo[1, tp] * (1 - sav)), 2) == c1
  #round((lcv * 5) + (ctabo[1, tp] * (1 - lcv)), 2) == c2
  print(all(v1 == (c1 - c2) / c1 * 100))
}  # all true

# Carbon bias by bins 
# DT-derived values
mu0a <- do.call(rbind, lapply(1:20, function(x) {
  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma0"))
}))
cb_stats$f1$sa30
cb_stats$f10$sa30[get("bin") == 19, ]

# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
                c("sa30", "f50", "sparse"))) {
  x <- chk[1] #names(i[[1]])[3]
  y <- chk[2] #names(i)[4]
  tp <- chk[3]
  print(paste(".", x, "..", y, "..", tp))
  d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
    #print(paste("......bin", j))
    d2 <- do.call(rbind, lapply(tp, function(z) {#:length(i[[y]][[x]]),..
      #print(paste("........variant", z))
      l3 <- abs(i[[y]][[x]][[z]])  # i[[y]][[x]][[z]]
      #print(paste("........vegetation:", names(l3)))
      d3 <- do.call(rbind, lapply(1, function(k) { #:length(l3)...
        rs <- gti_bins[[y]][[1]] == j #[[k]] == j    
        rs[rs == 0] <- NA
        o <- rs * l3 #[[k]]
        w <- area_wgts[[y]][[1]] * rs
        oo <- cbind.data.frame("lc" = rep(paste0(x, y, z), length(w)), 
                               getValues(o), getValues(w))
      }))
      d3f <- d3[!is.na(d3[, 2]), ]
    }))
    if(nrow(d2) > 0) { #(length(i[[x]][[y]]) * 10)) {
      qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                  weight.opts = list("weights" = d2[, 3])))
    } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
      qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
                                          na.rm = FALSE))
    } else {
      qs <- rep(NA, 7)
    }
  }))
  #print(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
  #          na.rm = TRUE)
  #print(all(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
  #          na.rm = TRUE))
  print(all(round(d1[, "mu"], 2) == 
              mu0a[V1 == x & V2 == tp, y, with = FALSE], na.rm = TRUE))
}
# all true

# another check of DT-derived stats against a raster-calculated variant
tst <- sapply(names(c_pct_diff)[1], function(x) {
  lapply(names(c_pct_diff[[1]]), function(y) {
    cellStats(c_pct_diff[[x]][[y]]$forest, mean)
  })
})
mean(round(unlist(tst)), 2) == mean(round(mu0[V2 == "forest", f1]), 2)

gti_gt0 <- Which(gti_agg$f1$gti > 0)
lc_gt0 <- calc(stack(lapply(lc_agg$f1, function(x) Which(x > 0))), sum)
all_gt0 <- gti_gt0 + lc_gt0
all_gt0 <- all_gt0 / all_gt0

filts <- sapply(names(c_pct_diff), function(x) {
  gti_gt0 <- Which(gti_agg[[x]]$gti > 0)
  lc_gt0 <- calc(stack(lapply(lc_agg[[x]], function(j) Which(j > 0))), sum)
  all_gt0 <- gti_gt0 + lc_gt0
  all_gt0 <- all_gt0 / all_gt0
})

tst <- sapply(names(c_pct_diff), function(x) {
  lapply(names(c_pct_diff[[x]]), function(y) {
    cellStats(filts[[x]] * c_pct_diff[[x]][[y]]$forest, mean)
  })
})

tstabs <- sapply(names(c_pct_diff), function(x) {
  lapply(names(c_pct_diff[[x]]), function(y) {
    cellStats(abs(filts[[x]] * c_pct_diff[[x]][[y]]$forest), mean)
  })
})

plot(c(1, 6), range(tst, tstabs), pch = "", xlab = )
for(i in 1:nrow(tst)) lines(1:6, tst[i, ], lty = 2)
for(i in 1:nrow(tstabs)) lines(1:6, tstabs[i, ], col = "red")

cellStats(all_gt0 * c_pct_diff$f1$sa30$sparse, mean)
cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, median)
cellStats(c_pct_diff$f1$sa30$forest, mean)
cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, mean)

gti_

```