### 2015-08-20 Rmd/et-bias.Rmd
ET bias analysis added (et-bias.Rmd), nearly complete. Updated abm district selection also, changed to 2011 dataset rather than 2007.  Redid figure for supplemental

---
title: "ET bias"
author: "Lyndon Estes"
date: "August 17, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Overview

Bias in PET estimates as calculated using different version of the cropland datasets to determine the vegetation properties in VIC.

# Analysis
## Prepare datasets

Bring in cropland datasets, ET grids, reproject the latter to SA Albers, mask, convert to monthly total ET, etc. 
```{r, eval = FALSE}
library(SAcropland)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
p_et <- full_path(p_root, "external/ext_data/vic/")

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
anms <- c("gti", snms)
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100)

# check
# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- 25
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(gti, stack(lc_list)), sumna)
namask2 <- !is.na(namask)  # set NAs to zero
# # namaskt <- raster(full_path(p_data, "namask.tif")) # load in NA mask
# # namaskt2 <- !is.na(namaskt)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum); cellStats(namask2t, sum)
# cellStats(namaskt, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts$f25, function(x) {
  r <- x[[1]]
  projection(r) <- sa.shp@proj4string
  r
})  # reshape a bit into list for subsequent use
names(awgts) <- "f25"
mask25k <- awgts$f25 > 0  # mask for et datasets

# read in ET estimates, but drop first year because of VIC problem in January
etf <- dir(p_et, pattern = "nc", full.names = TRUE)
etb <- lapply(etf, function(x) brick(x))
etb <- lapply(etb, function(x) dropLayer(x, i = 1:12))

# reproject them to Albers
etba <- lapply(etb, function(x) {
  b <- projectRaster(x, awgts$f25)
})

# rename layers and convert to total month mm
dts <- seq.Date(as.Date("1981/1/1"), as.Date("2008/12/31"), "days")
allmos <- substr(dts, 6, 7)
allyrs <- substr(dts, 1, 4)
ndays <- unlist(lapply(unique(allyrs), function(x) {
  unname(sapply(unique(allmos), function(y) {
    length(which(allyrs == x & allmos == y))
  }))
}))  # n days in each month in time series

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

# check that mm/month multiplies through correctly
tst <- etba[[1]] * ndays
plot(tst[[15]] - etba[[1]][[15]] * ndays[15])

# convert to monthly total mm
etba <- lapply(etba, function(x) x * ndays)
# plot(etba[[1]][[1]])

# year indices
yrs <- gsub("*.*_", "", lnames)
mos <- gsub("_.*", "", lnames)
yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))
mi <- sapply(unique(mos), function(x) which(mos == x))
```
<a href="#top">Back to top</a>

## Process ET time series

Figure out which months produce maximum ET values, calculate monthly means, overall country-wide time series, etc. 
```{r, eval=FALSE}
# Calculate mean, median, mode of dates in time series when PET max occurs
et_max <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    which.max(x[[ind]])
  }))
  s <- stack(calc(bi, max), calc(bi, modal), calc(bi, mean), calc(bi, median))
  names(s) <- c("maxmax", "mode", "mean", "median")
  s
})

# plot to see which statistic picks out PET peak date most effectively
plot(et_max$gti)  # all the same, not coherent
plot(et_max$gti %in% 6:9)  # median looks most sensible and coherent 
plot(et_max$gti - et_max$sa30)  # some differences in date of max
plot(et_max$gti - et_max$globmu) 
plot(et_max$gti - et_max$modmu)  
plot(et_max$gti - et_max$geow)

# Monthly mean values
et_momu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:ncol(mi), function(y) {
    ind <- mi[, y]
    calc(x[[ind]], mean)
  }))
  names(bi) <- unique(mos)
  bi
})

# do a focal max to separate out noisy areas
et_medsm <- focal(et_max$gti$median, w = matrix(1, 3, 3), fun = modal, pad = 3)
buf <- (!is.na(et_max$gti$median) - !is.na(et_medsm)) * et_max$gti$median
et_medsm[is.na(et_medsm)] <- 0
et_medsm <- buf + et_medsm  # fill back in areas edged out by moving window
# plot((et_medsm > 0) - (et_momu$gti[[1]] > 0))

# 1 month before and after peak
et_meds_m1 <- et_medsm - 1
# plot(et_medsm - 1)  # a few zero areas
et_meds_m1[et_meds_m1 < 1] <- 12  # set to 12 (December) if median - 1 = 0
et_meds_p1 <- et_medsm + 1
# plot(et_medsm + 1)  # nothing greater than 12

# calculate 3 month peak ET
gs_peak <- lapply(et_momu, function(x) {
  et1 <- stackSelect(x, et_meds_m1)
  et2 <- stackSelect(x, et_medsm)
  et3 <- stackSelect(x, et_meds_p1)
  stack(et1, et2, et3)
})

# Annual mean values
et_amu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
  et_ts <- cellStats(bi, mean)
  names(et_ts) <- 1981:2008
  list("gridmu" = calc(bi, mean), "cntry_ts" = et_ts)
})

# plot annual mean values
rbcol <- rainbow(4)
plot(1981:2008, et_amu$gti$cntry_ts, type = "l", las = 2)
for(i in 2:5) lines(1981:2008, et_amu[[i]]$cntry_ts, col = rbcol[i])

# figure out min_max years
mm_yr <- sapply(list(which.min, which.max), function(x) x(et_amu$gti$cntry_ts))

# annual totals
et_mutot <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
})
et_mm <- lapply(et_mutot, function(x) x[[mm_yr]])
```
<a href="#top">Back to top</a>

## Calculate ET differences
```{r, eval=FALSE}
# difference rasters and convert to percent
# between time series annual mean
ilist <- lapply(et_amu[2:5], function(x) x[[1]])
a <- et_amu$gti$gridmu
amu_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series mins
ilist <- lapply(et_mm[2:5], function(x) x[[1]])
a <- et_mm$gti[[1]]
mumin_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series maxs
ilist <- lapply(et_mm[2:5], function(x) x[[2]])
a <- et_mm$gti[[2]]
mumax_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between mean peak growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) calc(x, sum))
a <- calc(gs_peak$gti, sum)
gs3_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between median max growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) x[[2]])
a <- gs_peak$gti[[2]]
gsmx_diff <- lapply(ilist, function(b) {
  bl <- (a - b) / a * 100
})
```
<a href="#top">Back to top</a>

## Plot difference maps
### Reshape and disaggregate lists
```{r, eval = FALSE}
# function to reshape lists for use in disaggregate_rast_list
disaggl <- function(rl, lev = "f25") {
  bo <- lapply(rl, function(x) {
    b <- list(x)
    names(b) <- lev
    b
  })
  bo  
}

# reshape in list
l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
l2dag <- lapply(l2dag, disaggl)

# disaggregate
l2dagd <- lapply(l2dag, function(x) {
  disaggregate_rast_list(snms, "f25", x, namask)
})

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
```
<a href="#top">Back to top</a>

### Plot maps
```{r, eval=FALSE}
# calculate single set of breakranges for all sets
statsl <- lapply(l2dag, dstats)
rng <- do.call(cbind, lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
}))
rngdt <- data.table(t(rng))
rng <- unname(unlist(rngdt[, lapply(.SD, function(x) x[which.max(abs(x))])]))

fnm <- "et_bias_map.pdf"
legtext <- rep("%", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}

brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
cols[c(1, length(cols))] <- c("darkred", "purple3")
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(4, 5), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(j in 1:length(l2dagd)) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(snms[k])
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    if(j == 1) mtext(mcap[k], side = 2, line = 1, cex = cx)
    if(k == 1) mtext(ecap[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(-1, -0.5)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()
```
<a href="#top">Back to top</a>

## Calculate bias statistics

### Mean percent bias in ET estimates
```{r, eval=FALSE}
lev_vec <- "f25"
# create mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lcu <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev_vec

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg[[2]], function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})
names(bins) <- lev_vec

# weighted mean functions
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# calculate bias stats
vnms <- names(l2dag)
bstats <- lapply(vnms, function(x) {
  a <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wm, "mu", "bias", TRUE)
  b <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wma, "mua", "bias", TRUE)
  stat <- rbind(a, b)
})
names(bstats) <- vnms

# checks to see if stats correct
# bias_stats* functions correct
chk <- bias_stats(bins$f25, awgts$f25, lcu$f25$globmu, l2dag$mu$globmu$f25, wm,
                  "mu", "bias", TRUE)
all(chk[bvals == "mu", bias] == bstats$mu[il == "globmu" & bvals == "mu", bias])
# compared to raster
for(i in snms) {
  msk <- lcu$f25[[i]]
  msk[msk == 0] <- NA
  a <- l2dag$mu[[i]]$f25 * msk
  b <- awgts$f25 * msk
  d <- weighted.mean(values(a), values(b), na.rm = TRUE)
  print(round(d, 2) == bstats$mu[il == i & bvals == "mu" & bin == "all", bias])
}  # check

# extract stats
mu <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mu", "bias", x)
})
mua <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mua", "bias", x)
})

# reshape
mus <- rbindlist(lapply(names(mu), function(x) cbind(x, mu[[x]])))
mus[, ol := NULL]
setnames(mus, c("x", "il"), c("ETvar", "dataset"))
setkeyv(mus, c("dataset", "ETvar"))
muas <- rbindlist(lapply(names(mua), function(x) cbind(x, mua[[x]])))
muas[, ol := NULL]
setnames(muas, c("x", "il", "bias"), c("ETvar", "dataset", "absbias"))
setkeyv(muas, c("dataset", "ETvar"))
mutab <- mus[muas]
mutab[, lapply(.SD, function(x) max(abs(x))), .SDcols = 3:4]

mutab_xtab <- xtable::xtable(mutab, digits = 1)
print(mutab_xtab, type = "latex", 
      file = "paper/figures/et-bias.tex")
```
<a href="#top">Back to top</a>

The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.43%, so not much here.

### Bias in relation to rainfall

We'll end by having a look at the relationship between bias and rainfall

```{r, eval = FALSE}
# rainfall layer prepared for SA crop subsidy analysis
load("external/ext_data/climate/mag_dist_clim.rda")
pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
projection(presa) <- p4s  
presa <- projectRaster(presa, mask25k)  # project to Albers
presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask

plot(presa)
plot(abs(l2dag$mu$globmu$f25))
plot(presa, abs(l2dag$mu$globmu$f25))

```