### 2015-08-20 Rmd/magisterial-district-ABM-select.Rmd

---
title: "Districts for ABM"
author: "Lyndon Estes"
date: "June 26, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Overview

Selection of magisterial districts and associated landcover sets for use in ABM example.

# Data
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# landcover data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- raster(paths[2])  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
lc_list <- gti - dlist_1km

# magisterial district data cropland statistics (2011, but only to look at for 
# selection 
load("external/ext_data/gti_md.rda")
mu <- data.frame(t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(mu) <- c("N", "na", "f")

# district shapes, filtered by N missing data, minimum fraction, and area
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
mdsel <- as.numeric(row.names(mdfilt))

# Examine, including in QGIS
plot(md)
plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
         layer = "md_abms", driver = "SQLite")

```

# Selection

Looking at the data in QGIS, including some rainfall maps from earlier papers and the inter-tubes, it looks like the following districts are a good match:

   + Clocolan (269)
   + Ficksburg (274)
   + Fouriesburg (275)
   + Marquard (288)

```{r, eval=FALSE}
abm_md <- md[mdsel[mdsel %in% c(269, 274:275, 288)], ]
writeOGR(abm_md, dsn = "external/ext_data/md4.sqlite", layer = "md4", 
         driver = "SQLite")

# Deprecated: earlier used when running 2007 rather than 2011
# Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
# interest's sake
# gti_abm <- extract(gti, abm_md)
# abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
#   c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
# })))
# colnames(abm_mu) <- c("N", "na", "f")
# abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
 # -1.0  0.0 -0.4 -0.2  # tinf differences!
```

## Extraction
```{r, eval=FALSE}
gti4 <- crop(gti, extent(abm_md))
lc4 <- crop(lc_list, extent(abm_md))
cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif", 
            overwrite = TRUE)
```

## Plots
```{r, eval = FALSE}
brks1 <- seq(0, 100, 10)
cols1 <- rev(terrain.colors(length(brks1) - 1))
brks2 <- c(-100, -80, -60, -40, -20, -10, -5, -1)
brks2 <- c(brks2, rev(abs(brks2)))
cols2 <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks2) - 1)
cols2[c(1, length(cols2))] <- c("darkred", "purple3")
cx <- 1.1
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

# layout.show()
# pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
par(mfrow = c(3, 2), mar = c(0, 0, 0, 0), oma = c(2, 0, 0, 0))
plot(md, col = "grey", border = "transparent")
plot(abm_md, col = "red", border = "grey30", add = TRUE)
par(mar = c(3, 0, 0, 0))
plot(abm_md)
plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, add = TRUE, 
     legend = FALSE)
plot(abm_md, add = TRUE)
for(i in 2:5) {
  plot(abm_md)
  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks, col = cols, 
       add = TRUE, legend = FALSE)
  plot(abm_md, add = TRUE, lwd = 2)
  mtext(mcap[i - 1], side = 1, line = -4, adj = 0.7)
}
nct <- length(brks1) - 1
flex_legend(ncuts = nct, legend.text = "% cropland", legend.vals = rep("", nct),
            longdims = c(0.6, 0.9), shortdims = c(0.72, 0.012), colvec = cols1, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 3), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.6, 0.9, nct, constEWorNS = 0.72, resEWorNS = 0.02)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks1, srt = 270, adj = c(-0.15, 0.5), cex = 1.1)
nct <- length(brks2) - 1
flex_legend(ncuts = nct, legend.text = "% bias", legend.vals = rep("", nct),
            longdims = c(0.2, 0.8), shortdims = c(0.05, 0.015), colvec = cols2, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 6), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.2, 0.8, nct, constEWorNS = 0.047, resEWorNS = 0.012)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks2, srt = 270, adj = c(-0.1, 0.5), cex = 1.1)
dev.off()
```

![figure](../paper/figures/abm-selected-districts.png)