### 2015-08-27 Rmd/misc.Rmd

---
title: "Miscallaneous calculations"
author: "Lyndon Estes"
date: "August 24, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

Various small calculations made for the analysis.

```{r, eval=FALSE}
library(rgdal)
library(SAcropland)
library(rgeos)

p_dat <- full_path(full_path(proj_root("SAcropland"), "SAcropland"), 
                   "external/ext_data")

```

Area of sub-Saharan Africa and South Africa's share of that area
```{r, eval=FALSE}
eco <- readOGR(dsn = full_path(p_dat, "africa_ecofloristic_zones.sqlite"), 
               layer = "africa_ecofloristic_zones")
eco@proj4string <- CRS(projection(raster(nrow = 1, ncol = 1)))
af <- readOGR(dsn = full_path(p_dat, "africa_countries_alb.sqlite"), 
               layer = "africa_countries_alb")
af@data <- af@data[, c(1, 3:4)]
ecoalb <- spTransform(eco, af@proj4string)

# cut down countries
nms <- af$cntry_name[af$region == "Northern Africa"]
nms <- nms[nms != "Sudan"]  # keep Sudan
ssa <- af[!af$cntry_name %in% nms, ]
ssa <- ssa[-grep("Island|Sao|Verde|Mauritius|Helena|Comoros|Portugal|Seychell",
                 as.character(ssa@data$cntry_name)), ]
ssa <- gBuffer(ssa, width = 0)
# plot(ssa)

desert <- ecoalb[grep("desert", ecoalb$gez_term), ]
# plot(ecoalb[names(which.max(gArea(desert, byid = TRUE) / 10000 / 100)), ])
sahara <- as.numeric(names(which.max(gArea(desert, byid = TRUE))))

nosahara <- ecoalb[-sahara, ]
plot(nosahara, col = "red")

ovs <- which(!is.na(over(nosahara, ssa)))
ssaeco <- nosahara[ovs, ]
ssaeco <- ssaeco[-1, ]  # drop a couple Saharan mountains
ssaeco <- ssaeco[-1, ]
# plot(ssaeco[ssaeco@data$gez_term == "Tropical mountain system", ][1, ])
plot(ssa)
plot(ssaeco, col = "red", add = TRUE)

# calculate area of remainder
afarea <- gArea(ssaeco) / 10000
# plot(af[af$cntry_name == "South Africa", ])
saarea <- gArea(af[af$cntry_name == "South Africa", ]) / 10000
saarea / afarea

```

```{r, eval=FALSE}
a <- sample(1:100, size = 20)
b <- sample(1:80, size = 20)
mean(a - b)
sum(a) / sum(b)

```