### 2015-08-30 paper/figures/et-bias.tex

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Aug 19 16:18:01 2015
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & ETvar & dataset & bias & absbias \\ 
  \hline
1 & gs & geow & 0.2 & 0.8 \\ 
  2 & gspk & geow & 0.2 & 0.8 \\ 
  3 & mu & geow & 0.3 & 0.7 \\ 
  4 & mumn & geow & 0.3 & 0.6 \\ 
  5 & mumx & geow & 0.3 & 0.8 \\ 
  6 & gs & globmu & 0.1 & 1.2 \\ 
  7 & gspk & globmu & 0.2 & 1.4 \\ 
  8 & mu & globmu & -0.1 & 1.0 \\ 
  9 & mumn & globmu & -0.2 & 0.9 \\ 
  10 & mumx & globmu & 0.2 & 1.2 \\ 
  11 & gs & modmu & -0.5 & 0.9 \\ 
  12 & gspk & modmu & -0.4 & 1.0 \\ 
  13 & mu & modmu & -0.6 & 0.8 \\ 
  14 & mumn & modmu & -0.6 & 0.7 \\ 
  15 & mumx & modmu & -0.5 & 0.8 \\ 
  16 & gs & sa30 & 0.3 & 0.7 \\ 
  17 & gspk & sa30 & 0.2 & 0.8 \\ 
  18 & mu & sa30 & 0.5 & 0.6 \\ 
  19 & mumn & sa30 & 0.4 & 0.5 \\ 
  20 & mumx & sa30 & 0.4 & 0.6 \\ 
   \hline
\end{tabular}
\end{table}