### 2015-09-11 Rmd/agent-allocation-bias.Rmd
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

---
title: "Bias in Allocation of ABM Agents"
author: "Lyndon Estes"
date: "June 26, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Overview

Selection of magisterial districts and associated landcover sets for use in ABM example, and analysis of bias after initial allocation of agents performed based on available farmland estimated from 100 m downscaled versions of the landcover data.

# Data
```{r, eval=FALSE}
library(SAcropland)
library(raster)
library(rgdal)
library(lmisc)
library(readxl)
library(rgeos)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# landcover data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- raster(paths[2])  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
lc_list <- gti - dlist_1km

# magisterial district data cropland statistics (2011, but only to look at for 
# selection 
load("external/ext_data/gti_md.rda")
mu <- data.frame(t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(mu) <- c("N", "na", "f")

# district shapes, filtered by N missing data, minimum fraction, and area
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
mdsel <- as.numeric(row.names(mdfilt))

# Examine, including in QGIS
plot(md)
plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
         layer = "md_abms", driver = "SQLite")

```
<a href="#top">Back to top</a>

## Magisterial District selection

Looking at the data in QGIS, including some rainfall maps from earlier papers and the inter-tubes, it looks like the following districts are a good match:

   + Clocolan (269)
   + Ficksburg (274)
   + Fouriesburg (275)
   + Marquard (288)

```{r, eval=FALSE}
abm_md <- md[mdsel[mdsel %in% c(269, 274:275, 288)], ]
writeOGR(abm_md, dsn = "external/ext_data/md4.sqlite", layer = "md4", 
         driver = "SQLite")

# Deprecated: earlier used when running 2007 rather than 2011
# Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
# interest's sake
# gti_abm <- extract(gti, abm_md)
# abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
#   c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
# })))
# colnames(abm_mu) <- c("N", "na", "f")
# abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
 # -1.0  0.0 -0.4 -0.2  # tinf differences!
rgeos::gArea(abm_md, byid = TRUE) / 10000 / 100
```
<a href="#top">Back to top</a>

### Extract cropland percentage from selected districts
```{r, eval=FALSE}
gti4 <- crop(gti, extent(abm_md))
lc4 <- crop(lc_list, extent(abm_md))
cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif", 
            overwrite = TRUE)
cpct <- extract(gti4, abm_md)
dcrop_pct <- round(sapply(cpct, sum, na.rm = TRUE) / sapply(cpct, length), 1)
```
Cropland percentages:

  + Clocolan (269) - 45%
  + Ficksburg (274) - 39%
  + Fouriesburg (275) - 29%
  + Marquard (288) - 44%

<a href="#top">Back to top</a>

### Map selected districts, cropland percentage, and bias
```{r, eval = FALSE}
brks1 <- seq(0, 100, 10)
cols1 <- rev(terrain.colors(length(brks1) - 1))
brks2 <- c(-100, -80, -60, -40, -20, -10, -5, -1)
brks2 <- c(brks2, rev(abs(brks2)))
cols2 <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks2) - 1)
cols2[c(1, length(cols2))] <- c("darkred", "purple3")
cx <- 1.1
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

# layout.show()
pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
# png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
par(mfrow = c(3, 2), mar = c(0, 0, 0, 0), oma = c(2, 0, 0, 0))
plot(md, col = "grey", border = "transparent")
plot(abm_md, col = "red", border = "grey30", add = TRUE)
par(mar = c(3, 0, 0, 0))
plot(abm_md)
plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, 
     add = TRUE, legend = FALSE)
plot(abm_md, add = TRUE)
polygonsLabel(abm_md, labels = paste0(abm_md$name_2, " (", 1:4, ")"), 
              method = "buffer", cex = 0.9)
for(i in 2:5) {
  plot(abm_md)
  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks2, 
       col = cols2, add = TRUE, legend = FALSE)
  plot(abm_md, add = TRUE, lwd = 2)

  mtext(mcap[i - 1], side = 1, line = -4, adj = 0.7)
}
nct <- length(brks1) - 1
flex_legend(ncuts = nct, legend.text = "% cropland", legend.vals = rep("", nct),
            longdims = c(0.6, 0.9), shortdims = c(0.72, 0.012), colvec = cols1, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 3), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.6, 0.9, nct, constEWorNS = 0.72, resEWorNS = 0.02)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks1, srt = 270, adj = c(-0.15, 0.5), cex = 1.1)
nct <- length(brks2) - 1
flex_legend(ncuts = nct, legend.text = "% bias", legend.vals = rep("", nct),
            longdims = c(0.2, 0.8), shortdims = c(0.05, 0.015), colvec = cols2, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 6), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.2, 0.8, nct, constEWorNS = 0.047, resEWorNS = 0.012)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks2, srt = 270, adj = c(-0.1, 0.5), cex = 1.1)
dev.off()
```
<a href="#top">Back to top</a>

![figure](../paper/figures/abm-selected-districts.png)

<a href="#top">Back to top</a>

# Downscaling and agent allocation

Performed separately and described in the methods.

# Analysis of agent allocation bias
```{r, eval = FALSE}
abias <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx")))
abias <- abias[Alg == 1]
setkey(abias, MD)
hhs <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx"), 
                             sheet = 2))
setkey(hhs, MD)
abias <- abias[hhs]
abias

# calculate district level bias as percent
pctbias <- abias[var == "ag_area", 
                 lapply(list(sa30, globmu, modmu, geow), function(x) {
                   (gti - x) / gti * 100
                 })]

# HHs without land as percent total households
noland <- abias[var == "HHs_no_land", 
                lapply(list(sa30, globmu, modmu, geow), function(x) {
                  x / HHs * 100
                })]

# unallocated farmland
unalloc <- round(abias[var == "ag_area_unalloc", snms, with = FALSE] / 
  abias[var == "ag_area", snms, with = FALSE] * 100, 2)

cexs <- seq(1, 2, 1 / 3)
# pdf(full_path(p_fig, "agent-bias.pdf"), height = 5, width = 3)
pdf(full_path(p_fig, "agent-bias.pdf"), height = 3, width = 5)
cx <- c(0.7, 0.8, 0.6)
# png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
# par(mfrow = c(2, 1), mar = c(0.5, 2.5, 0.5, 1), mgp = c(1, 0.05, 0), 
#     tcl = -0.2, oma = c(3.5, 0, 0, 0))
par(mfrow = c(1, 2), mar = c(2, 2, 0.5, 0), mgp = c(0.75, 0.05, 0), 
    tcl = -0.2, oma = c(3.5, 0, 0, 0.1))
scols <- c("red", "orange3", "green4", "blue")
cols <- c(rep("red", 4), rep("orange3", 4), rep("green4", 4), rep("blue", 4))
plot(unlist(pctbias), unlist(noland), col = cols, cex = cexs, pch = 20, 
     ylab = "% HH without farms", xlab = "% cropland bias", cex.lab = cx[2], 
     cex.axis = cx[1])
plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
     ylab = "% land unallocated", xlab = "% cropland bias", cex.lab = cx[2], 
     cex.axis = cx[1])
par(xpd = NA)

yst <- seq(0.05, 0.23, 0.18 / 4)
xy <- cbind(grconvertX(rep(0.5, 5), from = "ndc"), 
            grconvertY(yst, from = "ndc"))
for(i in 2:length(yst)) {
  if(i == 5) {
    leg <- c("District", 1:4)
  } else {
    leg <- rep("", 4)
  }
  legend(x = xy[i, 1], y = xy[i, 2], legend = rep("",4), #legend = leg, 
         pch = 20, cex = cx[3], 
         horiz = TRUE, adj = c(4, -5.75),  
         col = scols[i - 1], pt.cex = cexs, bty = "n")#cex = cexs,)
}
yst <- seq(0.05, 0.23, 0.18 / 4)
xy <- cbind(grconvertX(rep(0.515, 5), from = "ndc"), 
            grconvertY(yst, from = "ndc"))
text(x = xy[, 1], y = xy[, 2], labels = c(mcap, "District"), 
     #srt = 90, #pos = 2, 
     #adj = c(-1, 1), 
     pos = 2, col = c(scols, "black"), cex = cx[1])
xst <- seq(0.545, 0.65, 0.105 / 3)
xy <- cbind(grconvertX(xst, from = "ndc"), 
            grconvertY(rep(yst[length(yst)], 4), from = "ndc"))
text(x = xy[, 1], y = xy[, 2], labels = 1:4, pos = 2, cex = cx[1])
dev.off()

```

<a href="#top">Back to top</a>

![figure](../paper/figures/agent-bias.png)

<a href="#top">Back to top</a>