### 2015-09-11 Rmd/et-bias.Rmd

---
title: "ET bias"
author: "Lyndon Estes"
date: "August 17, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Overview

Bias in PET estimates as calculated using different version of the cropland datasets to determine the vegetation properties in VIC.

# Analysis
## Prepare datasets

Bring in cropland datasets, ET grids, reproject the latter to SA Albers, mask, convert to monthly total ET, etc. 
```{r, eval = FALSE}
library(SAcropland)
library(RColorBrewer)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
p_et <- full_path(p_root, "external/ext_data/vic/")

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
anms <- c("gti", snms)
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100)

# check
# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- 25
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(gti, stack(lc_list)), sumna)
namask2 <- !is.na(namask)  # set NAs to zero
# # namaskt <- raster(full_path(p_data, "namask.tif")) # load in NA mask
# # namaskt2 <- !is.na(namaskt)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum); cellStats(namask2t, sum)
# cellStats(namaskt, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts$f25, function(x) {
  r <- x[[1]]
  projection(r) <- sa.shp@proj4string
  r
})  # reshape a bit into list for subsequent use
names(awgts) <- "f25"
mask25k <- awgts$f25 > 0  # mask for et datasets

# read in ET estimates, but drop first year because of VIC problem in January
etf <- dir(p_et, pattern = "nc", full.names = TRUE)
etb <- lapply(etf, function(x) brick(x))
etb <- lapply(etb, function(x) dropLayer(x, i = 1:12))

# reproject them to Albers
etba <- lapply(etb, function(x) {
  b <- projectRaster(x, awgts$f25)
})

# rename layers and convert to total month mm
dts <- seq.Date(as.Date("1981/1/1"), as.Date("2008/12/31"), "days")
allmos <- substr(dts, 6, 7)
allyrs <- substr(dts, 1, 4)
ndays <- unlist(lapply(unique(allyrs), function(x) {
  unname(sapply(unique(allmos), function(y) {
    length(which(allyrs == x & allmos == y))
  }))
}))  # n days in each month in time series

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

# check that mm/month multiplies through correctly
tst <- etba[[1]] * ndays
plot(tst[[15]] - etba[[1]][[15]] * ndays[15])

# convert to monthly total mm
etba <- lapply(etba, function(x) x * ndays)
# plot(etba[[1]][[1]])

# date indices
yrs <- gsub("*.*_", "", lnames)
mos <- gsub("_.*", "", lnames)
yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))  # year index
mi <- sapply(unique(mos), function(x) which(mos == x))  # month index
```
<a href="#top">Back to top</a>

## Process ET time series

Figure out which months produce maximum ET values, calculate monthly means, overall country-wide time series, etc. 
```{r, eval=FALSE}
# Calculate mean, median, mode of dates in time series when PET max occurs
et_max <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    which.max(x[[ind]])
  }))
  s <- stack(calc(bi, max), calc(bi, modal), calc(bi, mean), calc(bi, median))
  names(s) <- c("maxmax", "mode", "mean", "median")
  s
})

# plot to see which statistic picks out PET peak date most effectively
plot(et_max$gti)  # all the same, not coherent
plot(et_max$gti %in% 6:9)  # median looks most sensible and coherent 
plot(et_max$gti - et_max$sa30)  # some differences in date of max
plot(et_max$gti - et_max$globmu) 
plot(et_max$gti - et_max$modmu)  
plot(et_max$gti - et_max$geow)

# Monthly mean values
et_momu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:ncol(mi), function(y) {
    ind <- mi[, y]
    calc(x[[ind]], mean)
  }))
  names(bi) <- unique(mos)
  bi
})

# do a focal max to separate out noisy areas
et_medsm <- focal(et_max$gti$median, w = matrix(1, 3, 3), fun = modal, pad = 3)
buf <- (!is.na(et_max$gti$median) - !is.na(et_medsm)) * et_max$gti$median
et_medsm[is.na(et_medsm)] <- 0
et_medsm <- buf + et_medsm  # fill back in areas edged out by moving window
# plot((et_medsm > 0) - (et_momu$gti[[1]] > 0))

# 1 month before and after peak
et_meds_m1 <- et_medsm - 1
# plot(et_medsm - 1)  # a few zero areas
et_meds_m1[et_meds_m1 < 1] <- 12  # set to 12 (December) if median - 1 = 0
et_meds_p1 <- et_medsm + 1
# plot(et_medsm + 1)  # nothing greater than 12

# calculate 3 month peak ET
gs_peak <- lapply(et_momu, function(x) {
  et1 <- stackSelect(x, et_meds_m1)
  et2 <- stackSelect(x, et_medsm)
  et3 <- stackSelect(x, et_meds_p1)
  stack(et1, et2, et3)
})

# Annual mean values
et_amu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
  et_ts <- cellStats(bi, mean)
  names(et_ts) <- 1981:2008
  list("gridmu" = calc(bi, mean), "cntry_ts" = et_ts)
})

# plot annual mean values
rbcol <- rainbow(4)
plot(1981:2008, et_amu$gti$cntry_ts, type = "l", las = 2)
for(i in 2:5) lines(1981:2008, et_amu[[i]]$cntry_ts, col = rbcol[i])

# figure out min_max years
mm_yr <- sapply(list(which.min, which.max), function(x) x(et_amu$gti$cntry_ts))

# annual totals
et_mutot <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
})
et_mm <- lapply(et_mutot, function(x) x[[mm_yr]])
```
<a href="#top">Back to top</a>

## Calculate ET differences
```{r, eval=FALSE}
# difference rasters and convert to percent
# between time series annual mean
ilist <- lapply(et_amu[2:5], function(x) x[[1]])
a <- et_amu$gti$gridmu
amu_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series mins
ilist <- lapply(et_mm[2:5], function(x) x[[1]])
a <- et_mm$gti[[1]]
mumin_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series maxs
ilist <- lapply(et_mm[2:5], function(x) x[[2]])
a <- et_mm$gti[[2]]
mumax_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between mean peak growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) calc(x, sum))
a <- calc(gs_peak$gti, sum)
gs3_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between median max growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) x[[2]])
a <- gs_peak$gti[[2]]
gsmx_diff <- lapply(ilist, function(b) {
  bl <- (a - b) / a * 100
})
```
<a href="#top">Back to top</a>

## Plot difference maps
### Reshape and disaggregate lists
```{r, eval = FALSE}
# function to reshape lists for use in disaggregate_rast_list
disaggl <- function(rl, lev = "f25") {
  bo <- lapply(rl, function(x) {
    b <- list(x)
    names(b) <- lev
    b
  })
  bo  
}

# reshape in list
# l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
# names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff)
names(l2dag) <- c("mu", "mumx", "mumn", "gs")
l2dag <- lapply(l2dag, disaggl)

# disaggregate
l2dagd <- lapply(l2dag, function(x) {
  disaggregate_rast_list(snms, "f25", x, namask)
})

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
```
<a href="#top">Back to top</a>

### Plot maps
```{r, eval=FALSE}
# calculate single set of breakranges for all sets
statsl <- lapply(l2dag, dstats)
rng <- do.call(cbind, lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
}))
rngdt <- data.table(t(rng))
rng <- unname(unlist(rngdt[, lapply(.SD, function(x) x[which.max(abs(x))])]))

fnm <- "et_bias_map4.pdf"
legtext <- rep("%", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
# ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")
ecap <- c("Annual mean", "TS max", "TS min", "Peak")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}

# For Supplementals
# brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
brks <- c(-25, -15, -10, -5, -3, -1, 1, 3, 5, 10, 15, 25)
# cvec <- c("red", "orange", "grey80", "green", "blue4")
# cols <- colorRampPalette(cvec)(length(brks) - 1)
# cols[c(1, length(cols))] <- c("darkred", "purple3")
cols <- brewer.pal(length(brks) - 1, name = "RdBu")
cols[6] <- "grey80"
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(j in 1:length(l2dagd)) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(paste("...", snms[k]))
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    if(j == 1) mtext(mcap[k], side = 2, line = 1, cex = cx)
    if(k == 1) mtext(ecap[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(-1, -0.5)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()

fnm <- "et_bias_map.pdf"
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(2, 2), mar = c(0, 0, 0, 0), oma = c(4, 0, 1, 0))
for(j in 1) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(paste("...", snms[k]))
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    mtext(mcap[k], side = 3, line = -1, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.07, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.3, 0.3), c(-1, -0.75)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()

```
<a href="#top">Back to top</a>

## Calculate bias statistics

### Mean percent bias in ET estimates
```{r, eval=FALSE}
lev_vec <- "f25"
# create mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lcu <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev_vec

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg[[2]], function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})
names(bins) <- lev_vec

# weighted mean functions
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# calculate bias stats
vnms <- names(l2dag)
bstats <- lapply(vnms, function(x) {
  a <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wm, "mu", "bias", TRUE)
  b <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wma, "mua", "bias", TRUE)
  stat <- rbind(a, b)
})
names(bstats) <- vnms

# checks to see if stats correct
# bias_stats* functions correct
chk <- bias_stats(bins$f25, awgts$f25, lcu$f25$globmu, l2dag$mu$globmu$f25, wm,
                  "mu", "bias", TRUE)
all(chk[bvals == "mu", bias] == bstats$mu[il == "globmu" & bvals == "mu", bias])
# compared to raster
for(i in snms) {
  msk <- lcu$f25[[i]]
  msk[msk == 0] <- NA
  a <- l2dag$mu[[i]]$f25 * msk
  b <- awgts$f25 * msk
  d <- weighted.mean(values(a), values(b), na.rm = TRUE)
  print(round(d, 2) == bstats$mu[il == i & bvals == "mu" & bin == "all", bias])
}  # check

# extract stats
mu <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mu", "bias", x)
})
mua <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mua", "bias", x)
})

# reshape
mus <- rbindlist(lapply(names(mu), function(x) cbind(x, mu[[x]])))
mus[, ol := NULL]
setnames(mus, c("x", "il", "bias"), c("Variable", "Map", "Bias"))
setkeyv(mus, c("Map", "Variable"))
muas <- rbindlist(lapply(names(mua), function(x) cbind(x, mua[[x]])))
muas[, ol := NULL]
setnames(muas, c("x", "il", "bias"), c("Variable", "Map", "Abs Bias"))
setkeyv(muas, c("Map", "Variable"))
mutab <- mus[muas]
mutab[, lapply(.SD, function(x) max(abs(x))), .SDcols = 3:4]
setkey(mutab, "Variable")
vnmr <- c("Annual Mean", "29-year Max", "29-year Min", "Peak")
for(i in 1:length(vnms)) mutab[Variable == vnms[i], Variable := vnmr[i]]
setkey(mutab, "Map")
for(i in 1:length(snms)) mutab[Map == snms[i], Map := mcap[i]]
mutab[order(-rank(Map))]

caption <- paste("Mean actual and absolute biases (as \\%) for",              
                 "evapotranspiration variables derived from a 29-year time",
                 "series calculated by the VIC model, including the average", 
                 "total ET for the", 
                 "3 months of the year when ET is highest, the annual mean", 
                 "and the minimum and maximum annual ETs in the time series.")
mutab_xtab <- xtable::xtable(mutab, digits = 1, caption = caption)
print(mutab_xtab, type = "latex", file = "paper/figures/et-bias.tex",
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)
```
<a href="#top">Back to top</a>

The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.25%, so not much here.

### Bias in relation to rainfall

We'll end by having a look at the relationship between bias and rainfall

```{r, eval = FALSE}
# rainfall layer prepared for SA crop subsidy analysis
load("external/ext_data/climate/mag_dist_clim.rda")
pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
projection(presa) <- p4s  
presa <- projectRaster(presa, mask25k)  # project to Albers
presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask

plot(presa, breaks = seq(0, 1000, 100), col = topo.colors(9))
plot(abs(l2dag$mu$globmu$f25))
plot(presa, abs(l2dag$mu$globmu$f25))

```