### 2015-09-11 paper/figures/et-bias.tex

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Sep  9 13:32:11 2015
\begin{longtable}{llrr}
\caption{Mean actual and absolute biases (as \%) for evapotranspiration variables derived from a 29-year time series calculated by the VIC model, including the average total ET for the 3 months of the year when ET is highest, the annual mean and the minimum and maximum annual ETs in the time series.} \\ 
  \hline
Variable & Map & Bias & Abs Bias \\ 
  \hline
Peak & GeoWiki & 0.2 & 0.8 \\ 
  Annual Mean & GeoWiki & 0.3 & 0.7 \\ 
  29-year Min & GeoWiki & 0.3 & 0.6 \\ 
  29-year Max & GeoWiki & 0.3 & 0.8 \\ 
  Peak & GlobCover & 0.1 & 1.2 \\ 
  Annual Mean & GlobCover & -0.1 & 1.0 \\ 
  29-year Min & GlobCover & -0.2 & 0.9 \\ 
  29-year Max & GlobCover & 0.2 & 1.2 \\ 
  Peak & MODIS & -0.5 & 0.9 \\ 
  Annual Mean & MODIS & -0.6 & 0.8 \\ 
  29-year Min & MODIS & -0.6 & 0.7 \\ 
  29-year Max & MODIS & -0.5 & 0.8 \\ 
  Peak & SA-LC & 0.3 & 0.7 \\ 
  Annual Mean & SA-LC & 0.5 & 0.6 \\ 
  29-year Min & SA-LC & 0.4 & 0.5 \\ 
  29-year Max & SA-LC & 0.4 & 0.6 \\ 
   \hline
\hline
\end{longtable}