### 2015-09-24 Rmd/landcover-bias-plots.Rmd

---
title: "Plotting cropland biases"
author: "Lyndon Estes"
date: "06 October 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Plotting bias figures
# Spatial biases
## Datasets

```{r, eval = FALSE}
library(raster)
library(lmisc)
library(R.utils)
library(RColorBrewer)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
load(full_path(p_data, "bias_stats.rda"))  # bias statistics
load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
```

## Plots
Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

# disggregate selected rasters at selected levels
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- dlist_act[[x]][[y]]$g2011
    } else {
      r <- disaggregate(dlist_act[[x]][[y]]$g2011, 
                        fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms
```

## Plot of selected rasters and levels 
```{r, eval = FALSE}
# Plotting colors
# cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, -5, -1, 1, 5, 10, 25, 50, 75, 100.1)
cols <- c(rev(brewer.pal(length(brks) / 2, "Reds")[-1]), "grey80", 
          brewer.pal(length(brks) / 2, "Blues")[-1])
brklen <- length(brks) - 1
# cols <- cols(brklen)
# cols[c(1, length(cols))] <- c("darkred", "purple3")
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 7, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(3.7, 1.5, 1.5, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, breaks = brks,
         legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = -0.25, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = -0.5, cex = cx)
  }
}
flex_legend(ncuts = brklen, legend.text = legtext, 
            legend.vals = round(brks), longdims = c(0.2, 0.8), 
            shortdims = c(0.05, 0.01), colvec = cols, srt = c(270, 0), 
            horiz = TRUE, textside = "bottom", legend.pos = c(4, 7), 
            leg.adj = list(c(0.2, 0.35), c(0.5, -0.5)), cex.val = cx, 
            textcol = lcol, bordercol = lcol)
dev.off()
```
The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, and 100 km from South Africa's landcover product, the mean estimates from 
GlobCover and MODIS, and Geowiki/GLC-SHARE.

<p align="center"><img src="../paper/figures/bias_map.pdf" width="400px"></p>

<a href="#top">Back to top</a>

# Bias in relation to field density

For supplementary data.

## Bias plots at 1 km with boxplots

```{r, eval = FALSE}
#btype <- c("bias_act.pdf", "bias_abs.pdf")
# plot matrix
nc <- 4
m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, 
            ncol = 32, byrow = TRUE) 
cols <- c("red", "orange3", "green4", "blue", "yellow")
bcol <- c("grey50", "grey50", "grey50", "grey")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
fcol <- "grey"
lev <- names(dstats_all$act$sa30)
lvec <- seq(5, 100, 5)
pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
                       seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, #c(1, 0.5, 1, 0.5, 0.5, 0.5, 0.5, 1, 0.5, 1),
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                         rep("grey80", 10)),
           "yl" = list(c(-100, 100), c(0, 100)), 
           "yl2" = c(-20, 60), "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
                            " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           "ycol" = c("grey70"))  # plot parameters

pdf(full_path(p_fig, "biases_1km.pdf"), width = 7, height = 5)
lmat <- rbind(m, m + max(m))
l <- layout(lmat)
par(mar = c(3, 0, 0, 0), oma = c(1, 4, 2, 2))
for(bt in 1:2) {
  snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
  dset <- dstats_all[[bt]][snms]
  for(i in 1:length(snms)) { i
    d <- dset[[snms[i]]]$f1
    n <- nrow(d)
    plot(pp$x, pp$y, ylim = pp$yl[[bt]], xlim = pp$xl, type = "l", 
         lwd = pp$y1lwd[[bt]], axes = FALSE, 
         xlab = "", ylab = "")
    if(i == 1) axis(2, at = pp$yax1[[bt]], labels = pp$yax1[[bt]], las = 2, 
                    mgp = c(1, 0.6, 0))
    axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yaxp[[bt]], cex = 0.8)
    for(z in 1:length(pp$y2[[bt]])) {
      lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd =pp$y2lwd, col = pp$y2lc[[bt]][z])
    }
    for(k in 1:n) {
      a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
      if(!is.na(d[k, 2])) {
        boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
                  lwd = c(1, 0.5, 2))
      }
    }
    lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
    if(i == 1) mtext(ifelse(bt == 1, "Bias (%)", "Mean absolute error (%)"), 
                     side = 2, line = 2.2, cex = 0.8)
    if(bt == 1) mtext(lcnms[i], side = 3, line = 0, col = cols[i])
  }
}
mtext("% Cropland", side = 1, outer = TRUE, line = -1, cex = 0.8)
dev.off()
```

<a href="#top">Back to top</a>

## Bias with scale
```{r, eval=FALSE}
#lev <- lev[!lev %in% c("f5", "f300")]
levnms <- paste(gsub("f", "", lev), "km")
m2 <- unlist(lapply(seq(1, by = 1, length = length(lev) * 4), function(x) {
  rep(x, nc)
}))
m3 <- do.call(rbind, lapply(seq(1, length(m2), length(lev) * nc), function(x) {
  do.call(rbind, rep(list(m2[x:(x + length(lev) * nc - 1)]), nc))
}))

yl <- seq(pp$yl2[1], pp$yl2[2], 10)
ylcol <- c("black", "grey", "black", rep("grey", 5), "black")
lt <- c(5, 1)
ylwd <- c(1.5, 2)

pdf(full_path(p_fig, "biases_1-100km.pdf"), width = 7.5, height = 7)
l <- layout(m3)
# layout.show(l)
# i <- 1; j <- 1
par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 5, 2, 2))
for(i in 1:length(snms)) {
  highcounter <- rep(0, length(lev))
  for(j in 1:length(lev)) {
    plot(pp$x, pp$y, ylim = pp$yl2, xlim = pp$xl, type = "l", lwd = 1, 
         axes = FALSE, xlab = "", ylab = "", xaxs = "i")
    for(z in 1:length(yl)) {
      lines(pp$x, rep(yl[z], 2), lwd = pp$y2lwd, col = ylcol[z])
    }
    muv <- rep(NA, 2)
    for(bt in 1:2) {
      d <- dstats_all[[bt]][[i]][[j]]
      n <- nrow(d)
      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = ylwd[bt], 
            lty = lt[bt])
      muv[bt] <- round(mean(d[, 8], na.rm = TRUE))
      toohigh <- which(d[, "mu"] > pp$yl2[2])
      if(length(toohigh) > 0) {
        highcounter[j] <- 1
        xi <- toohigh[which.max(d[toohigh, "mu"])]
        text(xi, y = pp$yl2[2] + 4, labels = round(d[xi, "mu"]), cex = 1.1,
             col = "grey50", xpd = NA)
      }
    }
    text(5, pp$yl2[2] - 14, labels = paste(muv, collapse = " / "), cex = 1.1, 
         col = "grey10")
    if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
    if(j == 1) mtext(lcnms[i], side = 2, line = 3, cex = 1, col = cols[i])
    if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
    axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
         tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
    axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yl2[1])
    if(i == length(snms)) {
      axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
           tcl = -0.3, pos = pp$yl2[1])
    }
  }
}
mtext("% Bias/MAE", side = 2, line = 2, cex = 0.8, outer = TRUE, adj = 0.5)
mtext("% Cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
par(xpd = NA)
x <- grconvertX(0.75, from = "ndc", to = "user")
y <- grconvertY(0.07, from = "ndc", to = "user")
legend(x = x, y = y, legend = c("Bias", "Mean absolute error"), lty = lt, 
       lwd = ylwd, bty = "n", cex = 1.2)
dev.off()

```
<a href="#top">Back to top</a>

# Total bias

Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

# recreate cropland percentages
lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
# plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be

# calculate km2 cropland in LC datasets
lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007

# bias in total cropland relative to GTI
stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)

# N km^2 in country
#sapply(lc2007, function(x) cellStats(!is.na(x), sum))
# sapply(lc2011, function(x) cellStats(!is.na(x) , sum))
# sapply(gti, function(x) cellStats(!is.na(x) , sum))
totkm2 <- cellStats(!is.na(gti$g2011) , sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```
<a href="#top">Back to top</a>

Total square kilometers of South Africa being evaluated
```{r, echo = FALSE}
library(lmisc)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
load(full_path(p_data, "total-cropland.rda"))
totkm2
```

Total square kilometers of cropland according to 2007 and 2011 GTI data
```{r, echo = FALSE}
refkm2
```

Percent bias in total cropland area estimates relative to 2007 GTI data
```{r, echo = FALSE}
stats2007
```

And relative to 2011 GTI data
```{r, echo = FALSE}
stats2011
```

<a href="#top">Back to top</a>

```{r, eval=FALSE, echo=FALSE}
DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {
  dstats_all[[2]][[x]]$f1[, c(1, 8)]
}))
dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
plot(DF[, 1], DF[, 2])
lines(predict.lm(dlm, newdata = data.frame("V1" = 1:110)))
```