### 2015-09-24 paper/figures/totC-bias.tex

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Sep 23 22:19:49 2015
\begin{table}[ht]
\centering
\caption{Percent differences in total carbon stock estimates calculated from the reference maps and from each of the four cropland maps. Differences are evaluated for total carbon estimates either at the country scale or over just the agricultural regions (cropland $>$0.05\%), using the carbon densities of 5 different cover types to provide the values for the non-agricultural portions of each pixel (cover types indicated by column names).} 
\begin{tabular}{llrrrrr}
  \hline
Region & Map & Forest & Secondary & Shrubland & Grassland & Sparse \\ 
  \hline
Country & SA-LC & 2.6 & 2.5 & 2.5 & 0.1 & -2.1 \\ 
  Country & GlobCover & -2.1 & -2.0 & -2.0 & -0.1 & 1.7 \\ 
  Country & MODIS & -2.6 & -2.5 & -2.5 & -0.1 & 2.1 \\ 
  Country & GeoWiki & 0.6 & 0.5 & 0.5 & 0.0 & -0.5 \\ 
  Agricultural & SA-LC & -2.0 & -2.7 & -2.8 & -10.6 & -14.9 \\ 
  Agricultural & GlobCover & -161.9 & -156.3 & -155.5 & -95.9 & -63.6 \\ 
  Agricultural & MODIS & -1.6 & -0.8 & -0.7 & 8.4 & 13.3 \\ 
  Agricultural & GeoWiki & 7.7 & 7.3 & 7.2 & 2.9 & 0.5 \\ 
   \hline
\end{tabular}
\end{table}