### 2015-10-01 Rmd/magisterial-district-bias.Rmd

---
title: "District-level cropcover and bias"
author: "Lyndon Estes"
date: "June 4, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Bias as a function of typical crop cover

## Data
```{r, message=FALSE, results='hold'}
library(raster)
library(lmisc)
library(nlme)
library(mgcv)
library(ape)
library(fitdistrplus)
library(SAcropland)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data2, "ZAF_adm2.Rdata"))  # magisterial districts
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths)  # gti 2011
gti@crs <- sa.shp@proj4string
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Transform magisterial districts, remove Prince Edward Islands
md <- spTransform(gadm, CRSobj = sa.shp@proj4string)
md <- md[md$ID_2 != 313, ]
```

Mean area of magisterial district.
```{r}
round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
```

<a href="#top">Back to top</a>

# Analyses

## Extract data for magisterial districts

Both cropland percentages and absolute biases. 
```{r, eval=FALSE}
# GTI data
# mean cropland cover
gti_md <- extract(x = gti, y = md, progress = "text")
save(gti_md, file = "external/ext_data/gti_md.rda")
# load("external/ext_data/gti_md.rda")
gti_mu <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
}))

# mean no zeros
gti_mu0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x[x > 0], na.rm = TRUE), 1))
}))

# how many contribute when non-cropland removed, and which are non-zero
nonzero <- sapply(gti_md, function(x) length(x[!is.na(x) & (x > 0)]))
whichnonzero <- sapply(gti_md, function(x) which(!is.na(x) & (x > 0)))

# median
gti_median <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(median(x, na.rm = TRUE), 1))
}))

gti_median0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), 
    round(median(x[x > 0], na.rm = TRUE), 1))
}))

par(mar = c(0, 0, 0, 0))
plot(md[323, ])
plot(gti, add = TRUE)
plot(md[323, ], add = TRUE)

# Absolute biases at 1 km
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_abs1km <- stack(lapply(dlist_abs[snms], function(x) x$f1$g2011))
# cellStats(!is.na(dlist_abs1km), sum)
dlist_abs1km@crs <- sa.shp@proj4string
dlist_abs1km_md <- extract(x = dlist_abs1km, y = md, progress = "text")
save(dlist_abs1km_md, file = "external/ext_data/dlist_abs1km_md.rda")
# load("external/ext_data/dlist_abs1km_md.rda")

# check counts, primarily of NA values for each dataset
dlist_counts <- t(sapply(dlist_abs1km_md, function(x) {
  apply(x, 2, function(y) {
    c(length(y), length(which(is.na(y))))#, round(mean(y, na.rm = TRUE), 1))
  })
}))

chkfun <- function(x, tol = 0.001) {
  return(apply(x, 1, function(x) abs(max(x) - min(x)) < tol))
}
a <- seq(1, 7, 2)
b <- seq(2, 8, 2)
dlist_counts[which(!chkfun(dlist_counts[, a])), a]
dlist_counts[which(!chkfun(dlist_counts[, b])), b]

# bind stats
# MD
# some have more NAs than others, so select which LC set has max NAs
nas <- apply(dlist_counts[, seq(2, 8, 2)], 1, max)
all(gti_mu[, 1] == dlist_counts[, 1])

# then calculate mean per LC set per district
# mean
dlist_mu <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(colMeans(x, na.rm = TRUE), 1)
})))

# mean no zero
# dlist_mu0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
#   tab <- round(x)
#   round(apply(tab, 2, function(x) mean(x[x > 0], na.rm = TRUE)), 1)
# })))
dlist_mu0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
  tab <- round(dlist_abs1km_md[[x]])
  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
  round(apply(tab, 2, function(x) mean(x, na.rm = TRUE)), 1)
})))

# plot(dlist_mu0[, 2], dlist_mu02[, 2])
# i <- 2
# plot(gti_mu0[, 3], dlist_mu02[, i])
# plot(gti_mu0[, 3], dlist_mu0[, i])

# median
dlist_median <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(matrixStats::colMedians(x, na.rm = TRUE), 1)
})))
colnames(dlist_median) <- c("nas", snms)

# dlist_median0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
#   tab <- round(x)
#   round(apply(tab, 2, function(x) median(x[x > 0], na.rm = TRUE)), 1)
# })))
dlist_median0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
  tab <- round(dlist_abs1km_md[[x]])
  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
  round(apply(tab, 2, function(x) median(x, na.rm = TRUE)), 1)
})))

# bind function
bind_stats <- function(x, y, coords = xy, 
                       snms = c("sa30", "globmu", "modmu", "geow")) {
  dat <- cbind.data.frame(x, y)
  dat$nd <- apply(dat[, c(2, 4)], 1, max)  # max of GTI and LC NAs
  colnames(dat)[c(1, 3)] <- c("N", "gti")
  dat <- dat[, -2]
  dat <- cbind(coords, dat)
  dat$dmu4 <- round(rowMeans(dat[, snms]), 1)  # mean across all 4
  dat$dmu3 <- round(rowMeans(dat[, snms[snms != "globmu"]]), 1) # mu - globcov
  dat
}

# centroids for districts
xy <- as.data.frame(rgeos::gCentroid(md, byid = TRUE))  # get centroid coords

# bind to gti for comparisons 
mu <- bind_stats(x = gti_mu, y = dlist_mu)
mu$w <- mu$N - mu$nd  # weight - N contributing pixels
mu0 <- bind_stats(x = gti_mu0, y = dlist_mu0)
mu0$w <- nonzero  # weight - N contributing pixels (non-zeros)
med <- bind_stats(x = gti_median, y = dlist_median)
med$w <- med$N - med$nd  # weight - N contributing pixels
med0 <- bind_stats(x = gti_median0, y = dlist_median0)
med0$w <- nonzero  # weight - N contributing pixels (non-zeros)

```
<a href="#top">Back to top</a>

## Run regressions and check for spatial autocorrelation
```{r, eval=FALSE}
# check the polygons and rows line up for coordinates
# i <- 300
# a <- extract(x = gti, y = md[i, ], progress = "text")
# length(a[[1]])
# length(which(is.na(a[[1]])))
# mean(a[[1]], na.rm = TRUE)
# mu[i, ]

proc_df <- function(x, nm) {
  nms <- which(colnames(x) %in% c("x", "y", "gti", nm, "w"))
  dat <- x[, nms]
  colnames(dat)[3:4] <- c("pv", "dv")
  return(dat)
}

nms <- colnames(mu)[match(c(snms, "dmu4", "dmu3"), colnames(mu))]
lm_mu <- lapply(nms, function(x) {
  dat <- proc_df(x = mu, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_mu0 <- lapply(nms, function(x) {
  dat <- proc_df(x = mu0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med <- lapply(nms, function(x) {
  dat <- proc_df(x = med, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med0 <- lapply(nms, function(x) {
  dat <- proc_df(x = med0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})
```
<a href="#top">Back to top</a>

### Plot the fitted LM models
```{r, eval=FALSE}
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
cols <- c("red", "orange3", "green4", "blue", "black", "grey50")
lw <- c(rep(1, 4), 3, 3)
lt <- c(rep(2, 4), 1, 1)
xl <- c(0, 100)
yl <- c(0, round(max(mu[, snms] / 10 )) * 10)

# mean cropland in districts
pdf(full_path(p_fig, "biases_magdist.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu)) {
  dlm <- lm_mu[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# mean non-zero cropland
yl <- c(0, round(max(mu0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_no0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu0)) {
  dlm <- lm_mu0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median cropland
yl <- c(0, round(max(med[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med)) {
  dlm <- lm_med[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median of non-zero cropland
yl <- c(0, round(max(med0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med0)) {
  dlm <- lm_med0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()
```
<a href="#top">Back to top</a>

### Check for spatial autocorrelation in residuals
```{r, eval=FALSE}
dists <- as.matrix(dist(mu[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu[[i]]$model), dists_inv)$p.value)  
}

dists <- as.matrix(dist(lm_mu0[[1]]$dat[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu0[[i]]$model), dists_inv)$p.value)  
} 
```
<a href="#top">Back to top</a>

Present and significant. To deal with it, first try to use:

### A generalized least squares (GLS) with spatial autocovariance terms.

```{r, eval=FALSE}
# mean of all data
i <- "geow"  # geowiki
dat <- proc_df(x = mu, nm = i)
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains

# cropland only mean
dat <- proc_df(x = mu0, nm = i)
dat <- dat[!is.na(dat$pv), ]
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains
```

That seemed inadequate for removing autocorrelation in residuals. The GLS did not seem to properly deal with residual autocorrelation, probably because the spatial structure is complex, given the pattern of rainfall and cropland in it.  A better approach is to include coordinates directly in the model, leveraging the 2-D smoothing capabilities of the GAM model in the mgcv library.

### Use a GAM instead

First check fit of model family, and whether assumptions of normally distributed errors are met. 
```{r, eval=FALSE}
# stats.stackexchange.com/questions/99425/distribution-for-percentage-data
# quartz()
# dat <- mu0[, "sa30"]
dat <- mu[, "sa30"]
dat <- dat[!is.na(dat)]
dat[dat == 0] <- dat[dat == 0] + 0.1
descdist(dat, discrete = FALSE)  # check potential distributions

# Check what distributions work best on data (non-cropland removed)
dists <- c("norm", "lnorm", "exp", "cauchy", "gamma", "logis", "beta","weibull")
fit_distrs <- lapply(snms, function(x) {
  dat <- mu0[, x]
  dat <- dat[!is.na(dat)] / 100
  dat[dat == 0] <- dat[dat == 0] + 0.1
  fits <- lapply(dists, function(x) fitdist(dat, x))
  names(fits) <- dists
  fits
})
names(fit_distrs) <- snms
# lapply(fit_distrs, function(x) sort(sapply(x, function(y) y$aic)))
plot(fit_distrs[[2]]$lnorm)  # lognorm shows up as best fit

# Matters more on residuals from model fit, however. 
# dataset
dat <- proc_df(x = mu, nm = snms[4])  # interactively change response  
# dat <- proc_df(x = med0, nm = snms[1])  # and input data
dat <- dat[!is.na(dat$pv), ]
dat[, 4] <- dat[, 4] + 0.1  # add 1 to allow for log transform

# GAM assuming Gaussian
dgam <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            method = "ML")
summary(dgam)
AIC(dgam)
gam.check(dgam)
summary(fitdist(residuals(dgam), "norm"))
plot(fitdist(residuals(dgam), "norm"))

# GAM Gaussian with log link
dgam2 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = gaussian(link = "log"), method = "ML")
summary(dgam2)
AIC(dgam2)
gam.check(dgam2)
summary(fitdist(residuals(dgam2), "norm"))
plot(fitdist(residuals(dgam2), "norm"))

# GAM w/Gamma
dgam3 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = Gamma(link = "identity"), method = "ML")
summary(dgam3)
AIC(dgam3)  # bizarre large AIC
gam.check(dgam3)  # crazy deviance residuals
summary(fitdist(residuals(dgam3), "norm"))
plot(fitdist(residuals(dgam3), "norm"))

# GAM w/log
dgam4 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
             method = "ML")
summary(dgam4)
AIC(dgam4)
gam.check(dgam4)
summary(fitdist(residuals(dgam4), "norm"))
plot(fitdist(residuals(dgam4), "norm"))
# termplot(dgam4, terms = "poly(pv, 2)", ask = FALSE)
# pred <- predict(dgam4, newdata = newdat, type = "response", se.fit = TRUE)

# above, with smooth fit instead of quadratic
dgam4b <- gam(log(dv + 0.0001) ~ s(pv, k = 3) + s(x, y), weights = w, data = dat,
              method = "ML")
plot(dat$pv, dat$dv)
summary(dgam4b)
AIC(dgam4b)
gam.check(dgam4)
summary(fitdist(residuals(dgam4b), "norm"))
plot(fitdist(residuals(dgam4b), "norm"))

plot.gam(dgam4b, residuals = TRUE, select = 1, se = TRUE, all.terms = TRUE, 
         pages = 0)

```
<a href="#top">Back to top</a>

Testing each of the main landcover sets for model shape, the best fit in terms of AIC and assumption of normality of residuals is from a log-normal model.

## Plot final results

Helper functions
```{r, eval = FALSE}
predfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "response", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit)
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit),
                    lo = fit - (1.96 * pred$se.fit))
  return(pdat)
}

termfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "terms", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit[, 1] + coef(model)[1])
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit[, 1]),
                    lo = fit - (1.96 * pred$se.fit[, 1]))
  return(pdat)
}

polyfunc <- function(pdat) {
  coord <- cbind.data.frame("x" = c(rev(pdat$pv), pdat$pv), 
                            "y" = c(rev(pdat$lo), pdat$up))
}

```
<a href="#top">Back to top</a>

### GAM fits
```{r, eval = FALSE}
gam_mods <- lapply(nms, function(i) {
  dat1 <- proc_df(x = mu0, nm = i)
  dat1 <- dat1[!is.na(dat1$pv), ]
  dat1$dv <- dat1$dv + 0.1  # adding 0.1 to allow log transform
  dat2 <- proc_df(x = mu, nm = i)
  dat2 <- dat2[!is.na(dat2$pv), ]
  dat2$dv <- dat2$dv + 0.1  # adding 0.1 to allow log transform
  dgam1 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat1, 
               method = "ML")
  dgam2 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat2, 
               method = "ML")
  newdat <- with(dat1, 
                 data.frame(pv = 0:100, "x" = mean(dat1$x), "y" = mean(dat1$y)))
  pdat1 <- predfunc(dgam1, newdat)
  pdat2 <- predfunc(dgam2, newdat)
  # pdat1 <- termfunc(dgam1, newdat)
  # pdat2 <- termfunc(dgam2, newdat)
  return(list("mumod" = dgam1, "medmod" = dgam2, "pdat1" = pdat1, 
              "pdat2" = pdat2,  "data1" = dat1, "data2" = dat2))  
})
```
<a href="#top">Back to top</a>

### GAM plots
```{r, eval = FALSE}
# plot(xl, yl2, pch = "", ylab = "% bias)", xlab = "% cropland", xaxs = "i",
#      yaxs = "i", yaxt = "n")
# round(log(seq(1, 90, 10)), 2)
# exp(seq(0, 4.5, 0.5))

ptcols <- sapply(cols, function(x) makeTransparent(x, 80))  # CI colors
lw <- c(rep(2, 4), 3, 3)
yl1 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat1$fit) * 1.1)), 1))
yl2 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat2$fit) * 1.1)), 1))
pdf(full_path(p_fig, "biases_md_lnorm_gam_mu0.pdf"), width = 7, height = 4.5)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
plot(xl, yl1, pch = "", ylab = "Mean Absolute Error (%)", 
     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
     yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
# for(i in 1:length(gam_mods)) {
for(i in 1:4) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat1)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat1, col = cols[i], lwd = 1.5, lty = 1)
}
# legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
#        col = cols, bty = "n", xpd = NA)#, title = "Fit")
legend("topleft", mcap, lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

pdf(full_path(p_fig, "biases_md_lnorm_gam_mu.pdf"), width = 7, height = 4.5)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
plot(xl, yl2, pch = "", ylab = "Mean Absolute Error (%)", 
     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
     yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
#for(i in 1:length(gam_mods)) {
for(i in 1:4) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat2)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat2, col = cols[i], lwd = 1.5, lty = 1)
}
# legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
#        col = cols, bty = "n", xpd = NA)#, title = "Fit")
legend("topleft", mcap, lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

# Check p-values for all model terms
lapply(gam_mods, function(x) {
  m1 <- summary(x$mumod)
  m2 <- summary(x$medmod)
  rbind("mu" = round(c(m1$p.table[, 4], m1$s.pv), 3), 
        "med" = round(c(m2$p.table[, 4], m2$s.pv), 3))
})  # all terms significant at p < 0.001 except 2nd order term on GlobCov med
```
<a href="#top">Back to top</a>

### Final plot of magisterial districts, for supplementals
```{r, eval=FALSE}
provs <- unique(md$ID_1)
greys <- paste0("grey", seq(15, 95, 10))
pdf(full_path(p_fig, "md_map.pdf"), width = 7, height = 7, bg = "transparent")
par(mar = rep(0, 4))
plot(sa.shp, lty = 0)
for(i in 1:length(provs)) {
  plot(md[md$ID_1 == provs[i], ], col = greys[i], #border = "transparent", 
       add = TRUE, lwd = 1)
}
dev.off()

```
<a href="#top">Back to top</a>