### 2015-11-03 Rmd/overview.Rmd
Full draft completed, ready for some edits.

---
title: "Analysis of cropland biases"
author: "Lyndon Estes"
date: "15 November 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
---

## Overview

The following files contain the code used to conduct each portion of the analysis described in the accompanying paper __Landcover Data Limits Our Understanding of Global Change__.

#### [Preparing reference maps](cropland-preprocess.html)
#### [Downloading and preparing cropland maps](fetch-land-cover.html)
#### [Cropland map bias and accuracy](compare-landcover.html)
#### [Accuracy versus cropland density](magisterial-district-bias.html)
#### [Carbon bias and accuracy](carbon-bias.html)
#### [Evapotranspiration bias and accuracy](et-bias.html)
#### [Yield and production bias and accuracy](yield-bias.html)
#### [Agent-based model sensitivity](agent-allocation-bias.html)