### 2015-12-01 Rmd/c1-landcover-preprocess-1.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

---
title: "Pre-processing cropland datasets"
author: "Lyndon Estes"
date: "29 September 2014"
output:
  html_document:
    theme: spacelab
    toc: yes
---

# Pre-processing reference cropland data layer
<!-- This analysis was undertaken to determine the bias in satellite-derived landcover products at different
spatial scales, using GeoTerraImage's cropland maps for South Africa to assess this error.-->
The following describes the initial pre-processing of landcover datasets that was
undertaken for an analysis of error in landcover data.

## Gridding the SA landcover data
Much of the original analysis was undertaken on the Mapping Africa server, using postgis/postgres as well as pprepair to do much of the cleaning and conversion of the GTI (reference) dataset into a 1 km resolution cropland percentage cover dataset. The following steps were taken:

### Cleaning up fields in GTI data
This was done for both the 2007 and 2011 datasets, using a script that pulled out the fields, found the bad ones, read them into an R data object, ran pprepair over it, and then stuck the cleaned geometries back into a postgis table holding the full dataset. The intersection was then performed.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-postgis.R")  # run on mapper.princeton.edu
```

### Calculating field areas

The next step was running the following script to calculate the area for different cropcover classes in each 1 km^2^ pixel.  This created two core gridded datasets, cover2007.tif and cover2011.tif.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-analyze.R")  # run on mapper.princeton.edu
```