### 2015-12-01 Rmd/c3-cropland-1.Rmd

---
title: "Calculating cropland bias and accuracy"
author: "Lyndon Estes"
date: "01 October 2014"
output:
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
    toc_depth: 4
    number_sections: true
---

# Cropland errors, part 1
This section of the analysis calculates the differences between the reference cropland percentages and those from the various landcover products. It draws from results generated by the [cropland pre-processing](cropland-preprocess.html). Bias statistics calculated here are primarily for the supplementals.

## Analyses
```{r, eval = FALSE}
library(lmisc)
library(raster)
library(SAcropland)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

### Load cropland test grids 
Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing, so GLC-Share dropped)

```{r, eval = FALSE}
salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
ds <- full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)])
globsa_list_sum <- lapply(ds, raster)
fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
modsa_list_sum <- lapply(fnames, raster)
# glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
lclist <- c(salc, globsa_list_sum, modsa_list_sum, geow)  # into list
names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu",
                   "modmax", "geow")
```

#### Reference data
```{r, eval = FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
```
<a href="#top">Back to top</a>

#### Apply NA masks
```{r, eval = FALSE}
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(stack(gti), stack(lclist)), sumna)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(namask2, sum)
namask[namask == 0] <- NA
# cellStats(namask, sum)
writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
            overwrite = TRUE)

# mask out all joint NAs
gti <- lapply(gti, function(x) mask(x, namask))
lclist <- lapply(lclist, function(x) mask(x, namask))

```

### Calculate differences at different levels of aggregation
This step draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids. This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages. The lists have two levels of nesting, here the upper provides the resolution, the inner provides the different landcover datasets.

```{r, eval = FALSE}
# Aggregate rasters
fact <- c(5, 10, 25, 50, 100)
#fact <- 25

area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)  # cells/pixel
lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters

# Calculate differences
# actual
dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
                            gti_agg, lc_agg, "a - b", silent = FALSE)  
# absolute
dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  
dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
                            gti_agg, gti_agg, "a - b", silent = FALSE)

save(dlist_act, file = full_path(p_data, "d_grid_act.rda"))
save(dlist_abs, file = full_path(p_data, "d_grid_abs.rda"))

# Write out the gti cover raster and the actual bias rasters for running impact of bias examples in land
# surface model (25 km resolution)
r <- gti_agg$f25$g2011
projection(r) <- projection(salc)
writeRaster(r, filename = full_path(p_data, "gti_2011_25km.tif"))
b <- brick(stack(lapply(c("sa30", "globmu", "modmu", "geow"), function(x) {
  dlist_act[[x]]$f25$g2011
})))
projection(b) <- projection(salc)
writeRaster(b, filename = full_path(p_data, "lc_bias_25km.tif"))

# r <- raster(full_path(p_data, "gti_2011_25km.tif"))
# lcb <- brick(full_path(p_data, "lc_bias_25km.tif"))
# plot(r - lcb, axes = FALSE, box = FALSE)
# plot(lcb)
# plot(round(lc_agg$f25$geow - (r - lcb$lc_bias_25km.4), 7))

```
<a href="#top">Back to top</a>

### Error versus cropland density
#### Weighting rasters 
Based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
```{r, eval = FALSE}

# cropland cover bins
binv <- seq(0, 100, 5)
gti_bins <- lapply(gti_agg, function(x) {
  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
})
```

#### Calculate bias/MAE, pooling MODIS and GlobCover across all versions
Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
```{r, eval = FALSE}
# Set up indexing parameters
fact_all <- c(1, fact)# full aggregation levels
lev_vec <- paste("f", fact_all, sep = "") 
nm_rt <- c("sa30", "glob", "mod", "geow")

# Reshape the list of actual landcover differences
dlists <- lapply(list(dlist_act, dlist_abs), function(i) {
  resh <- lapply(nm_rt, function(x) {  # select sensor
    l2 <- lapply(lev_vec, function(y) {  # isolate by scale
      l3 <- lapply(i[grep(x, names(i))], function(z) z[[y]])  # sensor by scale
        named_out(l3, names(i)[grep(x, names(i))])
      })
    named_out(l2, lev_vec)
  })
  named_out(resh, nm_rt)
})
names(dlists) <- c("act", "abs")
rm(dlist_act, dlist_abs)  # remove original to make space

```

#### Then calculate bias/MAE across datasets

Note: this uses an older and slower version of the code. Subsequent analyses use the faster version based on data.tables.  
```{r, eval = FALSE}
dang <- Sys.time()
# i <- dlists[[1]]; x <- names(i)[1]; y <- lev_vec[5]; j <- 14; z <- 1; k <- 1
dstats_all <- lapply(dlists, function(i) {
  print("processing 1 of 2 main bias lists")
  dstats <- lapply(names(i), function(x) {
    print(paste("..", x))
    l1 <- lapply(lev_vec, function(y) {
      print(paste("....", y))
      d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
        print(paste("......bin", j))
        d2 <- do.call(rbind, lapply(1:length(i[[x]][[y]]), function(z) {
          print(paste("........variant", z))
          l3 <- i[[x]][[y]][[z]]
          d3 <- do.call(rbind, lapply(1:length(l3), function(k) {
            rs <- gti_bins[[y]][[k]] == j    
            rs[rs == 0] <- NA
            o <- rs * l3[[k]]
            w <- area_wgts[[y]][[1]] * rs
            oo <- cbind.data.frame("lc" = rep(names(i)[z], length(w)), 
                                   getValues(o), getValues(w))
          }))
          d3f <- d3[!is.na(d3[, 2]), ]
        }))
        if(nrow(d2) > (length(i[[x]][[y]]) * 10)) {
          qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                      weight.opts = list("weights" = d2[, 3])))
        } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
          qs <- c(nrow(d2), rep(NA, 5), 
                  Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], na.rm = FALSE))
        } else {
          qs <- rep(NA, 7)
        }
      }))
      cbind(binv[-1], d1)
    })
    named_out(l1, lev_vec)
  })
  named_out(dstats, names(i))
})
dut <- Sys.time() - dang  # 33 minutes
save(dstats_all, file = full_path(p_data, "bias_stats.rda"))
```

<a href="#top">Back to top</a>

### Differences between 2007 and 2011 reference
```{r, eval = FALSE}
dstats_gti <- lapply(lev_vec, function(x) {
    #x <- lev_vec[[2]]
    i <-  dlist_gti$g2007[[x]][[1]]
    print(paste("..", x))
    d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(y) {
      #y <- 1
      print(paste("....", y))
      rs <- gti_bins[[x]][[2]] == y    
      rs[rs == 0] <- NA
      o <- rs * i
      w <- area_wgts[[x]][[1]] * rs
      oo <- cbind.data.frame(rep(y, length(w)), getValues(o), getValues(w))
      oof <- oo[!is.na(oo[, 2]), ]
      if(nrow(oof) > 10) {
        qs <- c(nrow(oof), box_stats(oof[, 2], weighted = TRUE, 
                                     weight.opts = list("weights" = oof[, 3])))
      } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
        qs <- c(nrow(oof), rep(NA, 5),  
                Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
      } else {
        qs <- rep(NA, 7)
      }
    }))
    cbind(binv[-1], d1)
})
save(dstats_gti, file = full_path(p_data, "bias_stats_gti.rda"))

```
<a href="#top">Back to top</a>

Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e. to make sure rasters weren't incorrectly indexed: all fine)); box_stats with weighted quantile option compared to previous commit where wtd.quantile was used separately (fine). Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)

### 2011 reference-test bias/accuracy
Whole country and agricultural areas only
```{r, eval = FALSE}
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids

# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
lc_pct_diff <- lapply(dlist_act[snms], function(x) {
  sapply(x, function(y) list(y$g2011))
})

# construct agricultural area union mask
lev <- names(lc_pct_diff[[1]])
lcu <- lapply(lev, function(x) {
  print(paste(".", x))
  lcb <- lapply(snms, function(y) {
    print(paste("...", y))
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

bins <- lapply(gti_bins, function(x) x$g2011)
awgts <- lapply(area_wgts, function(x) x[[1]])

# calculate means for 2011
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
mu_dt <- rbind(a, b)
mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)  # ag area means
mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)  # ag area abs mean
mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)  # country mean
mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)  # country abs
# across bin mean and abs mean
mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
setnames(mu_bin, "V1", "bias")
mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
setnames(mua_bin, "V1", "bias")
# mu_dt[il == "sa30" & bin != "all" & ol == "f1" & bvals == "mua", mean(bias)]

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
bv <- "mua"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(lc_pct_diff[[x]][[y]])
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

lcf_out11 <- do.call(rbind, lapply(1:length(svec), function(i) {
  ao <- do.call(rbind, lapply(snms, function(x) {
    ai <- sapply(lev, function(y) {
      get(svec[i])[ol == y & il == x, bias]
    })
    named_out(ai, paste(c(1, fact), "km"))
  }))
  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
}))

caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
                 "relative to the 2011 reference map for each aggregation",
                 "scale calculated over the entire country, for the union of",
                 "agricultural regions (cropland $>$ 0), and as density", 
                 "independent means, wherein the mean bias/MAE values",
                 "for each of 20 cropland cover classes (representing",
                 "5\\% increments of cover 0\\% to 100\\% defined",
                 "by the reference map) were calculated and then averaged.")
lcfout_xtab <- xtable(lcf_out11, digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2011-bias.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

# do.call(rbind, lapply(snms, function(i) {
#   mu_dt[!like(bin, "all") & ol == "f1" & il == i & bvals == "mu.nm", N]  
# }))
mubias_2011 <- mu_dt
save(lcf_out11, mubias_2011, file = full_path(p_data, "lcf_out2011.rda"))

```

### 2007 reference-test bias/accuracy
Whole country and agricultural areas only
```{r, eval = FALSE}
# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
lc_pct_diff <- lapply(dlist_act[snms], function(x) {
  sapply(x, function(y) list(y$g2007))
})

# construct agricultural area union mask
lev <- names(lc_list[[1]])
lcu <- lapply(lev, function(x) {
  print(paste(".", x))
  lcb <- lapply(snms, function(y) {
    print(paste("...", y))
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

bins <- lapply(gti_bins, function(x) x$g2007)
awgts <- lapply(area_wgts, function(x) x[[1]])

# calculate means for 2007
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
mu_dt <- rbind(a, b)
mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)
mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)
mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)
mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)
# across bin mean and abs mean
mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
setnames(mu_bin, "V1", "bias")
mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
setnames(mua_bin, "V1", "bias")

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
bv <- "mua"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(lc_pct_diff[[x]][[y]])
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

lcf_out07 <- do.call(rbind, lapply(1:length(svec), function(i) {
  ao <- do.call(rbind, lapply(snms, function(x) {
    ai <- sapply(lev, function(y) {
      get(svec[i])[ol == y & il == x, bias]
    })
    named_out(ai, paste(c(1, fact), "km"))
  }))
  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
}))

caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
                 "relative to the 2007 reference map for each aggregation scale",
                 "calculated over the entire country, for the union of",
                 "agricultural regions, and as density", 
                 "independent means, wherein the mean bias/MAE values",
                 "for each of 20 cropland cover classes",
                 "(representing 5\\% increments of cover 0\\% to 100\\% defined",
                 "by the reference map) were calculated and then averaged.")
lcfout_xtab <- xtable(lcf_out07, digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2007-bias.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

```