### 2017-01-27 Rmd/c10-location-impact.Rmd
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

---
title: "Yield and production bias/accuracy"
author: "Lyndon Estes"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Impact of cropland errors on location

Evaluated in terms of how landcover map error impacts the ability to locate areas having a particular quantity of interest.

## Data

Derived carbon (section 6) and yield estimates (section 8).

```{r, eval=FALSE}
library(SAcropland)
library(xtable)
library(RColorBrewer)
library(spatstat)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
# load(full_path(p_data, "yield-bias.rda"))

# SA provinces
prov <- readOGR("external/ext_data/provinces.sqlite", layer = "provinces")
prov <- prov[prov$id_1 != 9, ]  # remove Princeton Edward Islands
prov[prov$id_1 == 10, "id_1"] <- 9 # reset WC to id 10
load("external/ext_data/MZshapes.Rdata")

# yield and carbon data
load("external/ext_data/yieldprod-1km.rda")
load("external/ext_data/carbon-tier1.rda")
load("external/ext_data/cropland-1km.rda")
```

## Differences in location

If you are interested in selecting areas of highest yield/production/carbon potential, how far off are you relative to "truth". Use top quintile/decline to assess this, and then calculate mean nearest neighbor distance between two pixels of each map.

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...cropland
crclass <- lapply(cropl, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...yield
yclass <- lapply(cyld_agg$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...production
pclass <- lapply(cpagg1$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...carbon
cdat <- c("gti" = gti_c$f1, lc_c$f1)  # combine C 1 km sets
cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
  msk <- x$forest < 99.6
  cmu <- calc(x, mean)
  raster::mask(cmu, msk, maskvalue = 0)
})
cclass <- lapply(cdatf, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# carbon efficiencies (C / yield)
effclass <- lapply(1:5, function(x) { # x <- 1
  cyld <- cyld_agg$f1[[x]]
  cyld[cyld == 0] <- NA
  cdatfc <- crop(cdatf[[x]], cyld)
  eff <- cdatfc / cyld  # carbon efficiency
  # eff[is.infinite(eff)] <- NA
  rq <- quantile(eff, c(0, 0.9, 1))
  types <- cut(eff, breaks = rq)
  types > 1
})
names(effclass) <- names(cclass)

```

### Calculate nearest-neighbor distances

Between GTI high value areas and each of the four LC-derived datasets. 
```{r, eval = FALSE}

# calculate distance between GTI and each other dataset's classes. 
# pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
# pms2b <- pms2a[which(rowSums(pms2a) == 2 & pms2a$Var1 == 1), ]
# pms2b$cl <- c("YC1", "YC2", "YC3", "YC4")
# rownames(pms2b) <- 1:4

# Define neighborhood window
sa_owin <- owin(xrange = bbox(sa.shp)[1, ], yrange = bbox(sa.shp)[2, ],
                unitname = "meter")
# which1 <- function(x) x == 1  # selector function for r to xy

mudists <- lapply(list(crclass, yclass, pclass, cclass, effclass), function(x) {
  # x <- cclass
  mudist <- sapply(2:5, function(y) {  # y <- 2
    print(paste0("...", y))
    a <- rasterToPoints(x$gti, fun = function(x) x == 1)
    b <- rasterToPoints(x[[y]], fun = function(x) x == 1)
    pps <- lapply(list(a, b), function(z) {
      ppp(x = z[, 1], y = z[, 2], window = sa_owin)
    })
    names(pps) <- c("gti", names(x)[y])
    reord <- sort(sapply(pps, function(i) i$n), decreasing = TRUE)
    ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
    mean(ppnn$dist) / 1000
  })
  names(mudist) <- names(x)[2:5]
  mudist
})  
names(mudists) <- c("cropland", "yield", "prod", "carb", "eff")

# plot(stack(yclass))
# plot(is.na(yclass[[1]]) - is.na(yclass[[2]]))
# plot(is.na(cyld_agg$f1$gti) - is.na(cyld_agg$f1$geow))

# output plot
cols <- c("red", "orange3", "green4", "blue")
pcol <- makeTransparent(cols, alpha = c(120))
pcol2 <- makeTransparent(cols, alpha = c(30))

plot(1:4, pch = 20, col = pcol)
bs <- do.call(rbind, mudists[c(1, 2, 4)])
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")

pdf(full_path(p_fig, "location_offset.pdf"), height = 3.5, width = 4, 
    bg = "transparent")
par(mar = c(2, 3.5, 1, 0), bg = "transparent")
bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol, pcol2), 
              las = 2, border = "transparent", xaxt = "n", ylab = "", 
              ylim = c(0, 18), yaxt = "n")
axis(1, at = colMeans(bp), labels = lcnms, lty = 0, cex.axis = 0.9, 
     mgp = c(2, 0.25, 0))
axis(2, at = seq(0, 18, 2), labels = seq(0, 18, 2), las = 2)
mtext("km", side = 2, line = 2.25)
legend(x = 12.5, y = 18, legend = rep("", 4), pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
legend(x = 12.5, y = 17.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
legend(x = 12.5, y = 16.5, legend = rep("", 4), pch = 15, adj = 0, col = pcol2, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
text(x = 13, y = 17, labels = "Cropland", pos = 2, cex= 0.8)
text(x = 13, y = 16.25, labels = "Yield", pos = 2, cex = 0.8)
text(x = 13, y = 15.5, labels = "Carbon", pos = 2, cex= 0.8)
dev.off()

```