### 2017-08-31 inst/paper/figures/cropadj-bias-accuracy.tex

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Aug 30 17:02:57 2017
\begin{longtable}{llrrrrrr}
\caption{Bias and mean absolute errors (MAE) in statistically constrained cropland maps across aggregation scales, weighted by density of cropland cover in the reference map. } \\ 
  \hline
Metric & Map & 1 km & 5 km & 10 km & 25 km & 50 km & 100 km \\ 
  \hline
Bias & GLC-Share & 9.7 & 1.1 & 0.6 & 0.4 & 0.5 & 0.1 \\ 
  Bias & GlobCover & 34.5 & 18.3 & 14.5 & 10.6 & 7.6 & 4.6 \\ 
  Bias & MODIS & 17.8 & 5.5 & 3.2 & 1.3 & 0.1 & -1.3 \\ 
  Bias & SA-LC & 6.6 & 2.7 & 2.1 & 1.6 & 1.1 & 0.6 \\ 
  Accuracy & GLC-Share & 23.8 & 12.6 & 9.4 & 6.8 & 4.8 & 3.0 \\ 
  Accuracy & GlobCover & 42.3 & 27.3 & 23.3 & 18.8 & 15.6 & 11.2 \\ 
  Accuracy & MODIS & 33.8 & 21.5 & 18.4 & 15.3 & 12.7 & 10.6 \\ 
  Accuracy & SA-LC & 11.4 & 6.0 & 4.7 & 3.7 & 2.8 & 1.9 \\ 
   \hline
\hline
\end{longtable}